### 2019-01-09 manuscript.Rmd
Initial commit

---
title: "NATIONALIST AND HOME FIELD BIAS IN SPORT PERFORMANCE EVALUATIONS"
subtitle: "An example from the ski jumping World Cup"
runhead: Judging biases in skijumping
author:
- affiliation: Department of Medical Statistics, University of Oslo
  name: Leiv Tore Salte Rønneberg
- affiliation: Department of Sociology and Human Geography, University of Oslo.
  name: Torkild Hovde Lyngstad
- affiliation: European University Institute
  name: Juho Härkönen
date: "14 september 2017"
keywords: nationalist bias, home field bias, skijumping, judges
abstract: "High stakes and prizes in international sports competitions necessitate efficiently identifying legitimate winners. Illegitimate biases of non-independent judges and home field advantages represent two major threats to the legitimacy of international competitions. Judges’ nationalities represent one source of bias, as they may prefer athletes of their own nation. Judges may also favour athletes competing within their home countries, creating another potential bias. Such biases, conscious or unconscious, may affect the outcomes of competitions. The literature offers no complete consensus on the magnitude of such biases, hence not whether they affect competition outcomes or not. In this paper, we shed light on these problems, using international ski jumping as an example. We draw on data from the FIS World Cup competitions in the 2006/2007 and 2007/2008 seasons, and use several fixed-effects based regression designs. Our results reaffirm suspicions of national bias and home field advantage in major ski jumping competitions, but also show their magnitude is too little to be of major relevance to competition outcomes. "
thanks: "The authors are grateful for financial support from the Småforsk programme at the University of Oslo, and to Marte Knutsson for research assistance in the early phase of establishing the data set.  At publication time, all code and material used to produce this paper (with the exception of the proprietary data) will be available. The authors are also grateful to Matti Nykänen for being an enduring source of inspiration in many aspects of our lives."
output:
  pdf_document:
    keep_tex: yes
    latex_engine: xelatex
    template: /home/torkildl/Dropbox/Setups/thl/rmd-xelatex-template-ms.tex
  html_document: default
bibliography: /home/torkildl/Dropbox/Literature/thl-main-library.bib
csl: /home/torkildl/Dropbox/Literature/csl-files/american-sociological-review.csl
anonymous: yes
fontsize: 12pt
geometry: margin=2cm
header-includes:
    - \usepackage{placeins}
    - \usepackage{dcolumn}
    - \usepackage{tabularx}
    - \usepackage{lscape}
    - \usepackage{caption}
fig_caption: false
fig.align: "center"
fig.width: 5
---

```{r analysis, error=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#
# Analysis code for skijumping analysis
#
library(haven)
library(stringr)
library(lfe)
library(stargazer)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggthemes)
library(here)
library(broom)
library(purrr)
library(pander)

# Read in data
jumps <- as.data.frame(read_dta("./data_prepared.dta")) %>%
    select(order(names(.))) %>%
    arrange(jumpid)
tbl_df(jumps)

namevars <- str_c("judge_name",as.character(seq(1:5)))
nationvars <- str_c("judge_nation",as.character(seq(1:5)))
ptvars <- str_c("judge",as.character(seq(1:5)))
gathervars <- c(namevars,nationvars,ptvars)

j <- jumps %>%
    select(jumpid,starts_with("judge"),-judgetotal) %>%
    gather(key="key", value="jinfo", judge_name1:judge_name5, judge_nation1:judge_nation5,judge1:judge5) %>%
    mutate(judgeno = as.numeric(str_sub(key,start = -1))) %>%
    mutate(jvar = str_sub(key,end = -2)) %>%
    select(-key) %>%
    spread(key = jvar, value = jinfo) %>%
    arrange(jumpid) %>%
    left_join(.,jumps,by ="jumpid") %>%
    .[, !names(.) %in% gathervars]
tbl_df(j)

# Correct judge names
pairs <- matrix(c("YERLY James","YERRLY James","ASPHAUG Kolbjoern","ASPHAUG Kolbjoer","DALLE AVE Sandro","DALLE Ave Sandro","EGLOFF Ernst","EGLOFF Emst","KARJALAINEN Pirjo","KARJALAINEN Pirvo","KUUSINEN Eero","KUUSINE Eero","LOENG Geir","LOENG Geir S.","LORENZ Michael","LORENTZ Michael","NYMAN Tom","NYMAN Torn","SCHMIEDER Jorg","SCHMIEDER Joerg","SCHMIEDERER Konrad","SCHIMIEDERER Konrad"), ncol=2, byrow=TRUE)
pairs[1,1]
for (x in 1:nrow(pairs)) {
    pair <- pairs[x,]
    j <- j %>% mutate(judge_name = ifelse(judge_name==pair[2],pair[1],judge_name))
}

# Various data preparation
d <- j %>% 
  # Correct erroneous nation coding
  mutate(judge_nation = ifelse(judge_nation=="JAP", "JPN", judge_nation)) %>%
  # Convert to factors
  mutate(judge_name = as.factor(judge_name)) %>%
  mutate(judge_nation = as.factor(judge_nation)) %>%
  mutate(country = as.factor(country)) %>%
  # Generate indicator of national bias
  mutate(natbias = ifelse(competitornation==judge_nation,1,0)) %>%
  # Generate indicator of own nation represented in judge panel 
  group_by(jumpid) %>%
  mutate(represented = max(natbias)) %>%
  ungroup %>%
  # and indicator of compansation bias by other judges
  mutate(compensating = represented - natbias) %>%
  # Generate home field advantage dummy
  mutate(athome = ifelse(competitornation==country,1,0)) %>%
  # Generate indicator of national biar per country
  mutate(countrybias = ifelse(natbias==1,as.character(judge_nation),"NON")) %>%
  mutate(countrybias = as.factor(countrybias)) %>%
  # Home field advantage variables
  mutate(athome = ifelse(competitornation==country,1,0)) %>%
  mutate(competitorname = as.factor(competitorname)) %>%
  mutate(competitionplace = as.factor(competitionplace)) %>%
  mutate(athseason = as.factor(str_c(competitorname,as.character(season)))) %>%
  # Drop unneeded variables
  select(everything())

#
# ESTIMATIONS: NATIONAL BIAS  MODELS
#
# Model 1, Naïve model of national bias
model1 <- lm(data = d,judge ~ natbias)
summary(model1)

# Model 2: introducing jump fixed effects
model2 <- felm(data = d, judge ~ natbias | jumpid)
summary(model2)

# Model 3: Introducing judge fixed effects
model3 <- felm(data = d, judge ~ natbias  | jumpid + judge_name)
summary(model3)

# Model 4: Does national bias vary among nations?
model4 <- felm(data = d, judge ~ relevel(countrybias,ref = "NON") | jumpid + judge_name)
summary(model4)

# 
# * HOME FIELD BIAS MODELS
# 
# Model 5: Length model
e <- filter(d, judgeno==1)
model5 <- felm(data=e, lenpts ~ athome + competitionplace | athseason)
summary(model5)

# Model 6: Style points model
model6 <- felm(data=d, judge ~ athome + competitionplace + judge_name + speed + length| athseason)
summary(model6)
```

# INTRODUCTION

Many sports use qualified experts to rank and evaluate performances. For example, ski jumping competitions have a panel of five judges who independently evaluate each performance. Similar systems are used in diving, figure skating, and numerous other sports. The experts may completely or partially decide the outcome of competitions. Expert judgement is highly valued, and being a judge in sports competitions carries major prestige.

A prerogative of the system is that experts judge performances in an unbiased and neutral manner. However, there is a large potential for biases in such judgements (Meyer and Booker 1991). Experts may, for example, overestimate performances by competitors from their own country, overvalue certain stylistic elements, or unconsciously adjust evaluations in response to spectator expectations. Such biases may then be consequential in determining the outcome of competitions. Expert judgements are used well beyond the sports domain, and the relevance of studying expert judgements may inform several important societal domains. Expert judgements are widely used, for example, when evaluating applicants to universities, evaluating research grant proposals (Marsh et al. 2008), and in assessing job applicants (cf. Fasang 2006; Bursell 2012). Understanding the extent of and processes underlying such biases is important for any attempts to remove their effects.

In the present paper, we set out to study two possible types of bias in evaluations of ski jumping performances: nationalistic bias and home-field advantage bias. Nationalistic bias refers to biases in which the judges (positively) evaluate contenders from their own country, whereas home-field advantage bias arises if judges favour contenders from the country in which the contest is held. To answer our questions, we use data from the 2006-07 and 2007-08 seasons of the Federation International de Ski (FIS) World Cup series of ski jumping contests.

We contribute to the literature on nationalistic and home field advantage biases in several ways. First, by analysing both nationalistic and bias home field advantages on the same data, we can compare different aspects and sources of such bias. Second, by separately analysing the social and physical ramifications of home advantages, we can provide additional evidence on the nature of home field advantages. Third, by comparing our results to those obtained with similar data but from the markedly different context of the Olympic games, a highest-stakes-contest [@zitzewitz_nationalism_2006], we can shed light on the potential for biased evaluations in different institutional and incentive contexts.

# A REVIEW OF THE LITERATURE ON SUBJECTIVE BIASES IN SPORTS

A relatively large literature explores the existence and magnitude of subjective biases in sports evaluations. These biases may affect results or other outcomes in favor of what is in principle irrelevant aspects of competitors or other actors in the sport. Research have documented biases that promotes profits for the league owner or organization, racial and ethnic biases, nationalistic biases, as well as home advantage biases.

Research on biases in basketball illustrates the breadth of the field. In their study of American professional basketball, Price, Remer and Stone [@price_subperfect_2012] found that the referees were biased towards calls that would increase profits for the league. More specifically, in turnovers, the referees tend to favour the home team, the losing team in a particular game, and the losing teams of the play-off series. In turn these biases can lead to increased ticket sales and ad-revenue due to more exciting, closer games; and ultimately even extend the playoff series itself, garnering more revenue. Even though these biases may operate towards profit-maximizing goals, the authors note that they need not result from any conscious behaviour on behalf of the referees.

Two related papers explore a potential racial bias among NBA referees, and its consequences for betting markets [@price_racial_2010, @larsen_racial_2008]. The authors note that there is a significant difference between fouls called when the athlete and referees are of the same race versus when they are not. The difference is large enough to be an important factor in determining game outcomes, making it a viable tactic to pursue in betting markets. Analysing data from two seasons, they find that when the majority of judges are white, betting on the team that has more minutes (expected) played by white players will beat the point spread over 50% of the time.

## Nationalistic biases

Nationalistic or cultural bias, where judges favour athletes from their home country or culturally similar ones in a systematic way would be most evident in certain nationalistic competitions such as world cups or Olympics. Evidence of these kinds of biases have been found in many different sports, including rugby [@page_evidence_2010] , muay thai [@myers_evidence_2006], figure skating [@yang_cultural_2006], gymnastics [@callahan_cultural_2016], and surfing [@sampaio_three_2012].

In their study of the 2000 Olympic diving competition, Emerson, Seltzer and Lin [@emerson_assessing_2009] found evidence of strong biases in some of the judges; both nationalistic ones, and against competitors from certain other nations. Based on their model, they re-estimated the results, removing the biases to obtain the objective quality of the dive, and isolated a case where the medals might have changed hands – had the judges been completely unbiased. Zitzewitz [-@zitzewitz_nationalism_2006] made a similar analysis of datasets from the 2002 Salt Lake City Winter Olympics and found evidence of nationalistic biases in both ski jumping and figure skating. Evidence of bloc voting and vote trading was also found in figure skating where Germany, USA, Canada and Italy are more likely to be negatively biased towards Russia, Ukraine, France and Poland, and vice versa [@zitzewitz_nationalism_2006]. Bloc voting in figure skating is nothing new, however. Seltzer and Glass [-@seltzer_international_1991], for example, documented consistent bloc voting along cold-war lines in addition to nationalistic biases in the Winter Olympic games held from 1968 till 1988.

Vote trading in the 2002 Olympics led to a revamp of the scoring system for figure skating after a French judge admitted to being pressured to vote in favour of the Russian pair over the Canadian pair in the pairs’ competition final – effectively handing them the gold medal. Allegedly this was part of a deal with the Russian team, where they would reciprocate in a later competition in favour of the French. This scandal was examined quantitatively by Lock and Lock [-@lock_statistical_2003] using a bootstrap technique to look for inconsistencies in the scoring. Looking at the correlation between judges’ rank (of the athlete) and the final actual rank (a high correlation indicating an accurate judge), they isolated one inconsistent judge, but it was not the French one. The French judge had the best correlation of all the judges in that event; in other words, she was the most spot-on in her scoring, and no evidence of wrongdoing was found.

## Home field advantages

Home field advantage is another type of bias. Such advantages, where some competitors benefit from advantages from competing on the home field, may arise in many different ways. Several of the mechanisms that may produce home field advantages are pscyhological sources of superior performances, while others involve judges overrating home field competitors. This point illustrates that the mere presence of a biases is not necessarily illegitimate. In order to improve scoring and fairness in sports, one must understand the mechanisms behind them.

Some research into the home-field advantage has attempted to move our understanding of the causal mechanisms behind  the advantages. Carmichael and Thomas [-@carmichael_homefield_2005] outline the four most common explanations of this phenomena using soccer as an example. The first is a familiarity explanation – the home team is more familiar with the field, the stadium, the locker room, and so forth, which puts the players more at ease in the situation, allowing them to concentrate on performing at their best. Second, traveling to another field could result in fatigue, invoke various stressors, and cause disruptions in pre-competition routines for the away team. All these factors may lead to inferior performances by the away team, relative to their likely performance at the home field. Third, specific rules may favour the home team, giving them the upper hand. Last, there could be a crowd effect. The size of the crowd at the home field might improve individual or team performance. It might also make the judge to be more lenient for the home team, through psychological mechanisms.

Page and Page [-@page_evidence_2010] attributes some of the crowd effect to crowd noise, noting that the judge is under both time pressure (making a correct decision in a short time) and social pressure (from the crowd). A study by Nevill et al. [-@nevill_influence_2002] showed that crowd noise was related to the home-field advantage. Setting up an experiment with one group of judges watching a football match with sound, and another group without, they found that the group watching with sound called 15.5% fewer fouls against the home team than the group who watched in silence. This could also indicate that judges use crowd noise as a guideline in their judging work. The crowd effect may also take another direction, as Anders and Rotthoff [-@anders_homefield_2014] find that football clubs with notoriously violent fans have a larger home-field advantage than peaceful ones. Here, the theory is that the prospect of fans-to-judge violence causes a more lenient judge.

Research on the home-field advantage has also been done in various winter sports. Balmer et al. [-@balmer_home_2001] analysed home-field advantages the winter Olympics from 1908 to 1998, and found some evidence for it in figure skating, freestyle skiing, ski jumping, alpine skiing and short track speed skating. The authors examined the relevance of three of the explanations listed above (familiarity with local conditions, time zones and crowd effect), and found evidence for home condition familiarity.  Conditional on the time zones travelled, travel time as such did not matter. Most evidence, however, was found for the crowd effect theory, as sports that were judged only subjectively had a bigger home-field advantage than sports in which assessments were based on both subjective and objective criteria.

### Hypotheses

We propose several hypotheses on biases in international ski jumping competitions, based on the literature reviewed above. On the basis of earlier findings, we propose a first hypothesis that there are nationalistic biases in ski jumping. Specifically, we expect there to be a positive bias in scores when judges score performances by athletes who are from the same nation as themselves.

Our second hypotheses states that the gross bias is partly explained by controlling for characteristics of the jump and athlete. The empirical implication of is that the magnitude of the bias will attenuate once all unobserved factors related to the jump are controlled.

A third hypothesis is that this bias may be partly due to cultural influences on stylistic elements, and that the strength of the bias is lower in countries with strong traditions for ski jumping and higher in countries where the sport is relatively less important.

For home field advantages, we propose that home field advantages are measurable and impacts athletes performances. We further distinguish between the social aspects of competing on the home field and the physical aspects of competing on the home field. We hypothesize that the physical home field advantage is less salient than the social home field advantage, and thus less important for performances.

# DATA AND METHODS

To test the hypotheses proposed above, we use data from the FIS World Cup competitions in ski jumping from the 2006/07 and 2007/08 seasons. The original data were downloaded from the FIS web site, coded, and collated into complete data set for analysis. Our data include detailed information on each jump in each competition, including the five individual judge scores from the jump along with the identity and nationality of each judge in each competition. We also know where the competition was held.

Our data are now about a decade old, an age that in many social science areas may raise questions about the value of our analysis. We have made efforts to obtain newer and larger-scale data from owners of data on skijumping competitions. Our requests made to both Federation International de Ski and the Norwegian Skiing Assocation were rejected for legal reasons. However, even if more recent data would be slighly preferable, we are not worried about the analysis' "expiry date". The reason is that the phenomena under study, nationalist and home field biases, are neither new nor old, but general in nature and similar phenomena are likely to persist in any kind of evaluation setting. Our analysis, we would argue, has relevance far beyond the scope of our data, the skijumping sport, and perhaps even beyond the sporting domain.

We exclude team events from both seasons, leaving us with a data set of 51 competitions and 3,772 individual jumps. Table 1 outlines the dimensionality of the data set. The total score of the jump is the sum of the length score and the stylistic score. In our data set, individual judge evaluations of each jump are available, as well as additional technical data on the length and speed of each contestant’s jumps.

```{r tab1, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
# TABLE 1. Descriptive statistic
descheading <- c("Statistic","Mean / Number"); numdigs = 2
thetable <- as.data.frame(rbind( 
 c("No. of competitor nations", length(unique(jumps$competitornation))),
 c("No. of unique skijumping hills", length(unique(jumps$competitionplace))),
 c("No. of unique competitors",length(unique(jumps$competitorname))),
 c("No. of unique judges", length(unique(d$judge_name))),
 c("No. of jumps", nrow(jumps)),
 c("No. of competitions", length(unique(jumps$competition))),
 c("Avg. judge score", as.numeric(round(mean(as.numeric(d$judge)),numdigs))),
 c("Avg. jump length", as.numeric(round(mean(d$length),numdigs))),
 c("Avg. points", as.numeric(round(mean(d$points),numdigs))), deparse.level=0 ))
colnames(thetable) <- descheading
pander(thetable)
```

The analysis proceeds in several steps. First, we estimate nationalistic biases using a sequence of regression models where more complexity is incorporated for each model. Then, we estimate home field advantage. Finally, we simulate how important nationalistic biases and home field advantages are in final rankings of the competitions.

## Estimation of nationalistic biases

Following the general approach used in most studies of nationalistic biases (in particular, Emerson et al. (2009) and Zitzewitz (2006)), we analyse systematic tendencies in scores by individual judges and by the judges’ nationality. We begin with a simple model and add complexity in subsequent steps to approach a stricter test of nationalistic bias. In the first, simple, model, a dummy variable $\phi(I=J)$ indicates whether the judge comes from the same country (J) as the athlete (I) or not. The model can be written as

\begin{equation}
s_{ijp} = B∙ \phi(I=J) + \epsilon_{ijp}
\end{equation}

where $s_{ijp}$ is the score given by judge $j$, on performance $p$, by a certain athlete $i$, and $\epsilon_{ijp}$ is the error term. The $B$ coefficient in this model estimates the difference in expected score if the athlete and the judge come from the same country.

However, such a model does not take into account relevant differences between nations. A Norwegian judge may give higher scores to a fellow countryman’s performance due to a correspondence of their definitions of valuable stylistic elements. For example, Norway and Finland enjoy long ski jumping traditions, whereas neighbouring Sweden, an otherwise similar country, does not. The importance of tradition became salient when Jan Boklöv, a Swede, pioneered the V-style in ski jumping but was consistently disadvantaged by the judges. His style was not comme-il-faut among judges and competitors in ski jumping at the time, and thus heavily penalized by judges.

Our next step is an attempt to remove such confounding factors, by exploiting the other judges’ evaluations of the same jump in estimation of nationalistic bias. Assuming that the other four judges give an unbiased score, a more precise way to estimate the bias is to add a fixed effects-term for each jump to the model (in effect including a dummy variable for each jump). In such a model, our estimator is based on individual judges’ score deviances from the average score for each jump. If the average score represents a more objective assessment of the performance, deviations from it can be interpreted as signs of bias. Our second model is therefore

\begin{equation}
s_{ijp} = B∙\phi(I=J) + q_p + \epsilon_{ijp}
\end{equation}

where $q_p$ is the jump fixed effect. This model will not confuse differences in stylistic culture with nationalistic biases, comparing the potentially biased judge against the four other judges for each jump. Following Zitzewitz (2006), we also estimated a model with so-called compensating biases, relaxing the assumption of objectivity with the four remaining judges. This model allows for the remaining judges to compensate a potentially biased judge, by lowering their scores when they know that one of the others will up his score. We did not find any significant compensating bias in our data, but this particular model did not converge and we thus abstain from reporting it in our set of results.

A limitation of the second model is that it does not capture any inherent leniency the judges may display. If a judge consistently scores performances higher or lower than other judges, then we cannot be sure that the observed bias is not just a particular judge always scoring lower or higher than the average judge. To control for this, we add a judge fixed effect, $l_j$, which leads to our third model:

\begin{equation}
s_{ijp} = B∙\phi(I=J) + q_p + l_j + \epsilon_{ijp}
\end{equation}

This model is an improvement over the previous ones, but limitations still remain. The model assumes that bias is constant for each judge across all judges from all nations. Ideally, we would want to allow for variations in bias between individual judges or groups of judges by interacting the judge fixed-effect with the indicator function for judge and competitor nation correspondence. This would in principle let us obtain individual bias estimates for each judge, but our data unfortunately do not allow for such a specific estimation. They do however allow for estimating an interaction between nationalistic bias and the home country of the judge. This leads to the model:

\begin{equation}
s_{ijp} = B∙(L_{j}∙\phi(I=J)) + q_p + \epsilon_{ijp}
\end{equation}

where $L_j$ is an added judge country fixed-effect. In this fourth model, the estimates for B provide an assessment of the average bias in scores by judges from different countries.

## Estimation of home field advantage

To analyse home field advantage we need to separate the problem of bias into the physical and the social domains. That is, whether an athlete actually performs better on the home field or whether he gets better scores by the judges due to social pressure or other factors. Thanks to the natural separation of style scores and length scores in ski jumping, this can be uncovered using similar models.

In order to observe a home field effect in points, be it length points or the judges’ style points, we have to compare scores on a home field with scores on away fields – leading to the assumption that a jumper is somewhat consistent in performance throughout a season. Previously, we used the four other judges as a more objective measure of performance quality. For home field advantage we have to assume that an athlete maintains a relatively constant quality of jumping throughout the season, using his away jumps as a guide to his general performance level. This can be a bold assumption, given that local weather conditions and individual athlete’s “daily form” also may affect these scores, as well as the tiny margins of timing required to “push off” at the right moment. Another problem is that countries with a strong tradition for ski jumping will have both stronger jumpers, host competitions more often due to existing infrastructure, and have more judges in the judge panel.

Again, we construct a dummy-variable using a function, $ω(I=H)$, which will take the value 1 if the jump takes place in country I, the athlete’s home country, and H being the competition country. We then wish to estimate the following model:

\begin{equation}
s_{irs}' = \~B' ∙ω(I=H) + a_{i} +λ_{r} +\epsilon_{ir}
\end{equation}

where $s_{irs}$ are the length points of the athlete, in a specific season, in a specific competition. $\~B'$ is the estimated home field effect in points and $a_i$ is is a fixed effect for the athlete-season combinations. The fixed effect also removes any effect of individual jumpers that “strong jumpers” from traditionally strong ski jumping nations by out estimating each athlete, in each season, separately. Lastly, taking into account the different sizes of the hills, which may result in a skewness that the K-point system fails to adjust for, we add a fixed effect for the hill represented by $λ_r$.

For the social part of the home field advantage we wish to estimate a similar model where style points awarded by the judges serves as the dependent variable.

\begin{equation}
s_{ijprs} = \~B ∙ ω(I=H) + a_{is} + λ_r + l_j + βx_p + \epsilon_{ijprs}
\end{equation}

In this model, we add fixed effects for judges as in (3) to capture leniency of individual judges. We cannot use jump fixed-effects to capture the inherent objective quality of the jump, because home advantage is invariant within a jump, but we can include two central jump characteristics in our data, namely length and speed. This is because “all else being equal, a ski jumper with the same take-off speed but better form while airborne will travel further” [@zitzewitz_nationalism_2006, 78]. We therefore add these two variables to the model, where they are represented by the x.

In accordance with norms of open social science, we offer our data and analysis log in an open repository at the time of publication.

# RESULTS AND DISCUSSION

## Nationalistic biases in judge scores

Our analysis of nationalistic bias proceeds in four steps. Table 2 presents the results from the first three estimations. The fluctuation in the number of observations is due to jumps with missing values on one or more of the independent variables, which have been deleted from the estimation.

\begin{landscape}
```{r tab2, echo=FALSE, results='asis'}
stargazer(model1, model2, model3, out.header=FALSE, header=FALSE,
    single.row=TRUE,  
    column.labels = c("Model 1: Naïve", "Model 2: Within-jump", "Model 3: Within-judge"),
    title="Results from regression models of national bias", 
    covariate.labels = c("National bias"),
    dep.var.caption  = "", dep.var.labels.include = FALSE,
    model.numbers = FALSE, model.names = FALSE,
    add.lines = c(" "),
    notes.align="l",
    notes = c("In models 2 and 3, the intercept is suppressed by the fixed effects estimation procedure."),
    notes.append=TRUE,
    df = TRUE, decimal.mark =".",
    ci=TRUE, ci.level=0.95, ci.separator = "  ",
    omit.stat=NULL,
    star.cutoffs = c(0.05,0.01,0.001)
    )
```
\end{landscape}

Our first model yields a positive and statistically significant estimate: judges award an average of 0.146 extra points to jumpers from their home country. This is approximately one-third of the minimum increment in judge scores in ski jumping (0.5). Compared with within-jump variance it only amounts to a quarter of a standard deviation. This result already suggests that the magnitude of nationalistic bias in judges’ scores is limited. The estimate may be biased by cultural and historical differences between countries that lead to an over- or underappreciation of certain styles. 	The second model takes up exactly this problem, and provides a within-jump estimate of the nationalistic bias. In this model, the coefficient is estimated at 0.068, markedly lower than in the first model. It is statistically significant at the 0.1% level. The magnitude is low, but there is clearly a tendency towards nationalistic bias.

The third model let us also include judge leniency, a measure of how strict a judge is in his or her evaluations. From this model, we obtain a similar estimate of nationalistic bias of 0.081. All estimates are well below 0.5, the minimum increment in judge scores used in ski jumping. Judges’ leniency display important variation, however, as the range of judges leniency coefficients is larger than 0.8 points. Figure 1 show this variation in a histogram.
\bigskip

```{r fig1, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=5, fig.align="center"}
judge_effects <- filter(as.data.frame(getfe(model3, se = FALSE)),fe=="judge_name")
figure1 <- ggplot(data=judge_effects, aes(x=effect)) + 
    geom_histogram(bins=15) +
    xlab("Leniency") + ylab("Number of judges") +
    theme_minimal() +
    ggtitle("Figure 1. Histogram of judges' leniency")  
figure1
```
\bigskip

The fourth and final model allowed estimating the variation in nationalistic biases for judges from different countries. The results are shown in Figure 2. Countries with fewer than 25 observations are not shown, as their estimates have large standard errors and thus are imprecisely estimated.  French judges appear most biased, followed by Polish, and Japanese judges. For the other countries, the estimated biases are fairly small and in several cases not statistically significant. It is difficult to provide a substantive interpretation of this variation in national biases, but it is nevertheless clear that the biases are relatively small in most cases.

\bigskip
```{r fig2, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=5, fig.align="center"}
nationbias <- broom::tidy(model4) %>% 
    mutate(country = str_sub(term, start=-3, end=-1)) %>% 
    filter(!(country %in% c("USA","SVK","SWE","ITA"))) %>%
    mutate(ocountry = factor(country, levels = .$country[rev(order(.$estimate))])) %>% 
    mutate(lowci = estimate-1.96*std.error, hici=estimate+1.96*std.error) %>% 
    arrange(estimate)

figure2 <- ggplot(data = nationbias,mapping = aes(x = ocountry, y = estimate)) +
    geom_point(size=3) +  
    geom_pointrange(aes(x=country, y=estimate, ymin = lowci, ymax=hici), fatten = 1.5) +
    #color="grey") +
    #geom_segment(aes(x=country, y=estimate, xend=country, yend=0)) +
    xlab("Nation") + ylab("Strength of bias") +
    theme_minimal() +
    ggtitle("Figure 2. Nation-specific estimates of judges' bias")
figure2
```
\bigskip

### Home field advantages

Table 3 presents the results for the two models, where we included effects of the hypothesized home field advantage. Following the distinction between the physical and social dimensions of such advantages, we specified one model for the actual length of the jump and another for style scores. The first column of Table 3 show the parameter estimates for the length model (Model 5). This model includes competitor-season-fixed effects, which implies that variations among different athletes and their form is taken into account and the estimate for differential jump lengths for home field competitors are purged of those influences. Our main interest lies with the dummy variable for the athlete performing in his home country. The parameter estimate for this dummy variable is not significantly different from zero. Athletes thus do not seem to perform better in their home countries, than when competing in other countries. The physical dimension of home field advantage is thus undetectable in our data.

\begin{landscape}
```{r tab3, echo=FALSE, results='asis'}
stargazer(model5, model6, out.header=FALSE, header=FALSE,
    single.row=TRUE,  omit = c("judge","competitionplace"), 
    column.labels = c("Model 5: Length points", "Model 6: Style points"),
    title="Results from regression models of home field advantage", 
    covariate.labels = c("Home field"),
    dep.var.caption  = "", dep.var.labels.include = FALSE,
    model.numbers = FALSE, model.names = FALSE,
    add.lines = c(" "),
    notes.align="l",
    notes = c("Both models include athlete and jumping hill fixed effects, and model 6 also include judge fixed-effects."),
    notes.append=TRUE,
    df = TRUE, decimal.mark =".",
    ci=TRUE, ci.level=0.95, ci.separator = "  ",
    omit.stat=NULL,
    star.cutoffs = c(0.05,0.01,0.001)
    )
```
\end{landscape}

We examine the social dimension of home field advantage in Model 6, where we include a parameter for bias in judge's scores if the athlete is competing in his or her home country. If there is home field advantage, this parameter should be measurably larger than zero. The estimate of this parameter is statistically significant, and larger than zero, suggesting that there is a social dimension of home field advantage in ski jumping. The effect is small and roughly of the same magnitude (0.074) as for nationalistic bias. The bias only amounts to a quarter of a standard deviation within each jump and it is well below the minimum increment for judge scores of 0.5 point. It seems unlikely that home field advantages is a serious threat to the validity of results of any competition.

# Our results in light of the extant literature

Our bias estimates were smaller, however, than for example those reported from the Salt Lake City 2002 Winter Olympics [@zitzewitz_nationalism_2006]. The finding of generally smaller effects than those previously reported is intriguing and point to a potentially important element affecting subjective evaluation of performances: when stakes are higher, as in the prestigious Olympic games, the potential influence of biased judgments is likewise increased. The reason our results show less biases than previous research may thus be that the stakes in each of the World Cup series events are lower than the stakes in e.g. Olympic events [cf. @zitzewitz_nationalism_2006]. The mechanisms that lead to biases are somehow more often invoked or invoked in a more influential version when stakes are higher. Another potential explanation points to individual judges’ career considerations, as objective performance in regular contests can later be rewarded by participation in higher prestige events, such as the Olympics. Our results thus complement Zitzewitz’s (2006) findings by pointing out situations in which biases are more (less) likely. Future research could compare competitions that differ in status and prizes, for example competitions at the regional and the national levels, and further test this idea of bias as emergent in high-stakes competitions.

Our estimates of nationalistic biases were also small enough that they would not alter the rankings in any of the contests. A comparison with the within-jump variance indicates that the bias only amounts to around a quarter of a standard deviation (cf. Model 3 in Table 2). In other words, the variation between the judges’ scores that is not due to nationalistic bias is much larger. If this variation is considered to be random deviations from a “true” score, it is clear that competitors should worry more about chance variations in judges’ evaluations (and possible idiosyncratic opinions) than about nationalist biases affecting the outcome of the competition.

We conducted separate analyses for home-field bias in performance (as indicated by the length of the jumps) and in judges’ evaluations of the stylistic elements in each jump. The results showed no effects on the jumpers’ actual performance, implying that the jumpers did not benefit from familiarity with the hill or from home-audience support in terms of jump length. Instead, home-field bias showed up in the judges evaluations of the jumps, even though the effect was here again of very small magnitude. Whether the bias stemmed from “crowd-pleasing”—which functions as a conflicting motivation to objectivity in the judges’ performance—or some other bias-generating mechanism is beyond the scope of this study.

As has been pointed out in previous research, biases in evaluations need not result from conscious actions by the judges. In related research on hiring practices in symphony orchestras [cf. @fasang_recruitment_2006], researchers have claimed that evaluators have non-conscious schemata based on external factors such as gender, which affect the perception of performances and which can be removed by blinded evaluations. In a similar manner, judge evaluations in sport contests can be affected by non-conscious biases. A judge may over- or underrate an athlete's performance based on her or his nationality, or the judge may use the behaviour of the audience as information on this performance. In such cases, ways to minimize the effects of such factors should be promoted to reduce the effects of biases on evaluations.

# CONCLUSION

We estimated nationalistic and home-field advantages in ski jumping using data from the entire FIS World Cup 2006/2007 and 2007/2008 seasons. From our results, we confirm parts of the extant literature, and document both national and home field biases in judges’ evaluations. The biases were generally small, and do not seem to pose a major threat to the validity of results from skijumping competitions. However, in league or series-type competition tournaments, the cumulative effects of natioanlistic bias may be important. Future research should emphasize this problem.

Biases in subjective evaluations of judges and referees may carry important implications for sports results, as exemplified by findings of potential effects on final rankings of athletes or on betting markets [@larsen_racial_2008, @price_racial_2010]. Given the historically political nature of sports—as exemplified in extreme cases by the boycotts of the United States and the Soviet Union of the Moscow and Los Angeles Olympics, respectively, the contested nature of the 1972 chess championships between Fisher and Spassky, and the many cases of violence between sports fans (and, at the most extreme end the alleged “Football War” between Honduras and El Salvador in 1969)—subjective biases in sports can even carry political implications. Knowledge of the sources of evaluative biases in competitive sports can help us better understand biases and discriminatory practices in other fields, such as student and employee recruitment and the allocation of research grants. Our findings of nationalistic and home-field biases have provided evidence that, in particular when assessed together with other related findings, add to our understanding on bias-generating practices. This knowledge may aid sport organizations, as well as other types of organizations, in their control of similar biases, and produce fairer, performance evaluations in the future.

# REFERENCES

### 2019-07-25 revised-manuscript/response_to_reviewers.Rmd
Commit before removing home advantage

---
title: "Response to EJSS reviewers' comments on manuscript #2019-0005"
output:
  pdf_document:
    keep_tex: yes
    latex_engine: xelatex
    template: /home/torkildl/Dropbox/Setups/thl/rmd-xelatex-template-memo.tex
  html_document: default
bibliography: /home/torkildl/Dropbox/Literature/thl-main-library.bib
csl: /home/torkildl/Dropbox/Literature/csl-files/american-sociological-review.csl
anonymous: yes
fontsize: 12pt
geometry: margin=2cm
header-includes:
    - \usepackage{placeins}
    - \usepackage{dcolumn}
    - \usepackage{tabularx}
    - \usepackage{lscape}
    - \usepackage{caption}
fig_caption: false
fig.align: "center"
fig.width: 5
---

We are grateful for being granted the opportunity to revise and resubmit our manuscript. We have considered all the comments by both the reviewers and the Editor. They have helped us craft a stronger manuscript than we otherwise would be able to do. We believe our responses to the comments and criticisms made have made the work a stronger contribution to the literature.

In what follows, we respond to each of the comments and explain how we accordingly have adapted our work. In the cases where we have chosen not to follow the advice, we motivate our choices for doing so.

Thank you for reading and commenting on our work!

# Responses to comments made by the Editor

From the document attached with the Editor's Decision letter, we identified twelve comments and criticisms. We list the items as we have read them below, in order of appearance in the document, along with a brief statement on how we have dealt with the issue identified by the Editor.

>*1. Overemphasis on the prestige of expert judgement and being a sports judge.*

We have reworded the period on prestige of sport judges.

>*2. The size of the scholarly literature*

We have rephrased the first period of the literature review.

>*3. Typo: Em-dash before "effectively handing them the gold medal"*

The period is rewritten in the new version.

>*4. Typo: "Our second *hypotheses*" should be "Our second hypothesis"*

This is corrected in the new version.

>*5. The age of the data used in the analysis.*

In the new version, we also use data from the 2015-2016 season's World Cup competitions. We describe how we obtained these data and how they were treated in detail under our responses to reviewer \#2's comments.

>*6. The argument that results from ski jumping competitions may have relevance beyond the sport, needs a more solid foundation and a more precise characterization of ski jumping.*

We have toned down this argument somewhat in the new version.

>*7. The Editor appreciates the thorough explanation of statistical models used.*

Thank you! In this version, we have only changed small bits of material to make the models even easier to understand.

>*8. Sources of national differences in tendencies of national bias (french vs. others) -- what may they be? The Editor suggests that the potential growth in the sport nationally may be a factor.*

We now build upon sociological theory on evaluation to propose hypotheses on variation in biases in the field of ski jumping.

>*9. Our estimates of bias are less than Zitzewitz', which may be the result of a general professionalization of the sport. New data may cast further light on this issue.*

We now, with our new data, explicitly discuss and test a hypothesis of change over time.

>*10. Editor suggest that safeguards against biases are already in place (5 judges from different nations; lowest and highest scores excluded).*

This is of course correct. In the new version, we make more room for discussing these safgeguards and their effects.

>*11. Editor suggests reference to sports-related political turmoil is too far fetched. Instead, we should provide more social context of the ski jumping sport.*

We have followed this advice, and removed the political turmoil-related reference.

>*12. Reference list: There were missing data for two references (Sampaio 2012 and Yang 2006).*

The appropriate information has been added for these references.

\smallskip

We are grateful to the Editor for her/his comments on our manuscript. They helped us writerethink and rewrite the work in a way that we believe is a stronger contribution to the field. Thank you!

# Responses to comments made by reviewer \#2

Let us first express our thanks for the Reviewer's thoughtful comments on our manusript. In the folling paragraphs we respond to each of the comments and criticisms. We quote the Reviewer's comments (with minor edits) first before offering our response.

> *Distortion effects on the evaluation of performance in ski jumping are to be investigated. One is the national origin of the  judges and the other is the home advantage of the athletes. The research situation on the problem of the influence of nationality seems to be quite well worked out. The following hypotheses are formulated (p. 6):*

> *1) There will be a "positive distortion" of the results when the judges evaluate performances of athletes who come from the same nation as themselves.*

> *2) Another hypothesis claims that it is precisely the assumed effect that diminishes in strength when other variables are controlled (here I wonder about the sense). It is normal and plausible that the more variables are put into a model, the lower the relative explanatory share of a single variable. It would be interesting if the explanatory value of a variable remained high despite the control of disturbing variables (secondary variance).*

Yes, it is typical that when a new variable (Z) is introduced the regression coefficient of a first variable (X) may change both magnitude and sign if the new variable is correlated with X. In our case, it may that competitors from certain nations are better on average than competitors from other nations, and it may be that the above-average nations also are more often represented on judge panels. This will induce a correlation between the same-nation indicator (X) and qualities of the performance (Z). Controlling for Z, should then reduce the magnitude of the regression coeffient for X.

Our estimate from model 1, the "naïve" model, may be overestimating the strength of bias for such reasons, and in model 2 we control for *all* unobserved characteristics of the jump that affects the judge scores.

> *3) Finally, in a third hypothesis, cultural influences on stylistic elements are assumed (and I wonder whether this would not be an explanation for the variance of the variable "national influence"). So why is this separated? The authors differentiate between nations with a strong tradition of ski jumping and those with a rather weak tradition. In a first step there are some points critical to mention:*

> *1) The authors use a fixed effects-term for each jump in their regression-model to remove confounding factors and thereby estimate national bias more precisely. This is done under the assumption that the other - in this case four - judges give an unbiased score. Although this often so called „true score" is the common modus operandi in terms of the questions to be answered, the authors are to be pointed out on the fact that this assumption only holds true if no set of five judges contains two or more judges of identical nationality. Specific rules from the regarded years concerning the composition of the judging panel in ski jumping could be presented to support the assumption in that sense.*

Yes, we use a fixed-effects term for each jump to remove confounding factors from characteristics of the jump. However, we believe this model does _not_ rely on an assumption of unbiasedness of the other four scores nor an assumption that there is only one judge from the competitor's nation in the judge panel. If there are multiple judges from the competitor's country, there will simply be multiple 1's among the five observations that make up the observations of judge scores for each jump. What identifies the B (in the revised ms denoted $\beta$) parameter in the models, is the within-jump *variation* in judge-competitor correspondence. The identification criterion is that the data must include  some performances (jumps) where there is one (or more) judges from the same nation as the competitor, and thus scores 1 on the dummy variable for national bias. As we have quite a few of such performances, we are able to identify and estimate an effect of nationalistic bias on judge scores.

> *2) Additionally the first part oft the third hypothesis proposed by the authors seems not to be testable the way it is formulated, it is rather formulated as an additional assumption.  („A third hypothesis is that this bias may be partly due to cultural influences on stylistic elements,[…]"). For the second part of the third hypothesis however, there was no test of the hypothesis performed, nor was the concept of differing countries according to tradition for ski jumping further elucidated.*

We agree with you that the third hypothesis was mangled and untestable the way it was formulated. In the revised version, we have rewritten the hypothesis section and motivated the hypotheses better, in accordance with the new theoretical background.

> *3) Furthermore on page 12 it is stated that there would be a fluctuation in the number of observations due to missing values in the independent variables. Table 2 yet shows 18.860 observations for all three models presented, which indicates that all the available observations (3772 jumps x 5 judges) have been used to calculate the first three estimations.*

Thank you for pointing this out. This was an error on our part. It is corrected in the current version of the manuscript.

> *Home advantage is divided into a physical effect and a social effect. These constructs should be explained more clearly (see e.g. p. 10). I suspect that - according to the state of research (referenced on p. 5) - actual factors influencing performance exist (the sporting performance in the home country is actually on average higher than in a host country), but also influenced due to social pressure from the audience on the adjudicators, which can be assessed as higher in the home countries of the athletes than in the host countries (this is also explained on p. 10, this would then again be a distortion of judgement).*

We have expanded our discussion of home field advantages somewhat in this version, and make amongst other things the discintion you point to in your comment. 
TODO: OR SHOULD WE JUST DROP HOME FIELD?

> *It is expectable that the more variables are put into a model, the lower the relative explanatory share of a single variable. It would be interesting if the explanatory value of a variable remained high despite the control of disturbing variables (secondary variance).*

> *Some remarks on the regression model: The following regression models are proposed.*

>*National affiliation of the judges:*
>* sijp = B-(Lj- ϕ(I = J)) + qp + ϵijp*
>* s = score*
>* B = coefficient*
>* Φ = Dummy variable (nationality judge and athlete equal or unequal)*
>* L = leniency ("mildness" of the judges)*
>* q = jump fixed effect (as far as I understood it is the "ski jump tradition" of a country, is also explained 
>* as dummy  variable, probably nations with ski jump tradition and those without such a tradition??, that 
>* should be explained better)*
>* ϵ = disturbance term*
>* (Unclear: where is the actual power to be measured)?*

As the Editor seemed to be satisfied with the explanation of the statistical models, we have only made smaller refinements to the text on this area. We are willing, of course, to provide even deeper explanations than we already give, if need be.

A brief comment: The "jump fixed effect" represents a dummy variable for each jump (each performance), which means that we "trend" the judge scores against their mean. This implies that the model is the "within-jump" estimator of nationalistic biases.

<!---
>* Home advantage (physical):*
>* s'irs = B'-( I = H) + ai + r + ϵijp*
>* (shouldn't it be ϕ'(I = H)? Then ϕ' would be a dummy variable for the match between the venue and the athlete's country of origin.)*
>* s' = points for the width (so no posture note; in the text on p. 11 the ' is missing at s!!,)*
>* a = "seasonal effect"?*
>* r = nature of the hill?*
>* Home advantage (social portion):*
>* sijprs = B-(I = H) + ais + r + lj + xp + ϵijp*
>* supplementary (to previous)*
>* s = again posture mark*
>* l = mildness of the judges*
>* x = should represent jump characteristics (width, speed ??)*

--->

>* Formal notes: The following sources do not appear in the bibliography: Meyer & Booker, 1991; Bursell 2012, both indicated on p. 2, are not listed in the bibliography.*

This has been corrected in the revised version.

>* Typos on p. 5 Line 5: pscyhological => psychological*

This is corrected in the revised version.

>* Overall I recommend major revisions. In general: The limitations of the study should be explained better.*

TODO: In the revised version, we have tried to make the limitations of our study clearer. To this end, we have added ... ...

Thank you for taking your time to read and commenting upon on our work. In our view that your comments made us produce a stronger manuscript than the original submission, and we hope you concur.

# Response to comments made by reviewer \#3

We respond the comments made by reviewer \#3 below.

>* The manuscript is a quantitative analysis of judges' biases and home field advantages on results in ski jumping, based on data from the FIS World Cup in 2006-2008. In general, the subject is relevant and interesting; and authors apply sound methodology and statistics. All formal academic standards are met, and the paper is well written. However, data are rather old, even though they could be easily complemented by more recent data.*

In this version, we use both our original data from the mid 2000s, but also a new data set. This data set was created by scraping the FIS website using automated procedures. In short, we wrote pieces of software that 
1. walked through the FIS website and identified URLs pointing to results in PDF files on FIS' server
2. downloaded all these PDFs automatically (and with time gaps to avoid overloading servers)
3. scraped results data from these files according to format specifications.

This procedure was very successful, and let us quickly assemble a rather large data set with results from the 2015-2016 season. The software will be available on the GitHub repository associated with this paper, once the paper is published.

The procedure also has limitations. We exclude some events due to differences in results formats, and include only the first page of results from each file. These minor issues should not affect the results of our statistical analysis in any way. The competitions we include should be considered a random sample of competitions from the season, as the format differences are in all likelihood completely exogenous to the actual results.

>*But most of all, insights are quite limited because the authors follow a solely empiricist approach. Theoretical considerations are missing from the start, and thus the empirical findings are not very intriguing in the end:*

We concur that the original submission was less rich in theory than it was in analysis. Our aim with the study is not to make a theoretial contribution above and beyond testing implications of existing theory drawn from across the social sciences. We have however, inspired by your comments, added relevant theoretical scaffolding for our analysis. In our opinion, this move has improved the mansucript over the original version.

>* The literature review is not very extensive and only comprises empirical studies. Hypotheses refer only broadly to some of the literature or seem to be generated ex-post, instead of being deduced from any theory. Not only sociological literature is missing, but there aren't any systematic theoretical considerations, e.g. on the role of judges, on nationalism in sports, or on the specific context of ski jumping in comparison to other sports. In the end then, authors face the typical problem of such an empiricist approach: Their results - significant but weak biases, which seem to be addressed quite well by the sport's regulations - do not "speak for themselves", and thus you wonder which insight you have gained. Usually, a more thorough theoretical consideration will also lead to a more insightful interpretation of the results. Instead of treating ski jumping as just one more arbitrary case and speculate about potential threats of biases (p. 20), authors should reconstruct the social context (or field, figuration, system) of ski jumping as a specific case. Then they would be able compare their findings to other sports and reflect upon the meaning of biases in different constellations.*

Inspired by your comments, we have rethought these issues and added some material on the practice of evaluations. Here, we sketch the ski jumping field as a field in a pseudo-Bourdieusian sense and connect the national variation in judge scores to struggles of definition of what good performances are. In particular, we build on the literature on evaluation that has flowed from the seminal work of Boltanski & Thevenon, Lamont and others. We also draw upon the historical example of the Boklöv style, which at first was not taken seriously but over time has come to dominate the field.

Thank you again for your efforts! Your comments made us rethink and add components to the analysis and manuscript that in turn led to a stronger contribution to the literature.

\bigskip

END.

### 2019-07-25 revised-manuscript/manuscript.Rmd
Commit before removing home advantage

--- 2019-01-09 manuscript.Rmd
@@
 In this paper, we shed light on these problems, using international ski jumping as an example.
-We draw on data from the FIS World Cup competitions in the 2006/2007 and 2007/2008 seasons, and use several fixed-effects based regression designs.
+We draw on data from the FIS World Cup competitions in the 2006-2008 and 2015-2016 seasons, and employ several fixed-effects based regression designs.
 Our results reaffirm suspicions of national bias and home field advantage in major ski jumping competitions, but also show their magnitude is too little to be of major relevance to competition outcomes.
-"
+There is a trend towards smaller nationalist biases  "
 thanks: "The authors are grateful for financial support from the Småforsk programme at the University of Oslo, and to Marte Knutsson for research assistance in the early phase of establishing the data set.
@@
 bibliography: /home/torkildl/Dropbox/Literature/thl-main-library.bib
-csl: /home/torkildl/Dropbox/Literature/csl-files/american-sociological-review.csl
 anonymous: yes
@@
 # Read in data
-jumps <- as.data.frame(read_dta("./data_prepared.dta")) %>%
+jumps <- as.data.frame(read_dta(here("data_prepared.dta"))) %>%
     select(order(names(.))) %>%
@@
 summary(model6)
+
+#
+# Season 2015-2016 analysis
+#
+season2016 <- data.table::fread(here("scrape/scraped-data-season-2016.csv"))
+
+scores16 <- season2016 %>% 
+    select(contains("judge"), competitorNation, country, place, speed, length, competitorName, gender) %>%
+    mutate(jumpid = row_number()) %>%
+    gather(key=jud, value=val, contains("judge")) %>%
+    mutate(info = case_when(str_detect(jud, "Name") ~ "judge_name",
+                            str_detect(jud,"Nation") ~ "judge_nation",
+                            TRUE ~ "judge")) %>%
+    mutate(ltr = str_sub(jud, start=-1)) %>%
+    select(-jud) %>%
+    group_by(jumpid, ltr) %>%
+    arrange(jumpid, ltr) %>% 
+    spread(key=info, value=val) %>%
+    ungroup %>% 
+    mutate(natbias = if_else(competitorNation==judge_nation,1,0)) %>%
+    mutate_at(c("judge"), as.numeric) %>% 
+    mutate(countrybias = factor(ifelse(natbias==1,as.character(judge_nation),"NON"))) %>% 
+    mutate(competitionplace = str_sub(place, start=1,end=-6)) %>% 
+    mutate(athome = if_else(country==competitorNation,1,0))
+
+# NATIONAL BIAS MODELS WITH 2016 DATA
+#
+### Model 1b: Naïve model
+model1b <- lm(data=scores16, judge ~ natbias)
+summary(model1b)
+
+### Model 2b: Introducing jump fixed effects
+model2b <- felm(data = scores16, judge ~ natbias | jumpid)
+summary(model2b)
+
+### Model 3b: Introducing judge fixed effects
+model3b <- felm(data = scores16, judge ~ natbias  | jumpid + judge_name)
+summary(model3b)
+
+# Model 4b: Does national bias vary among nations?
+model4b <- felm(data = scores16, judge ~ relevel(countrybias,ref = "NON") | jumpid + judge_name)
+summary(model4b)
+
+# HOME FIELD BIAS MODELS WITH 2016 DATA
+# 
+home16 <- season2016 %>% 
+    select(competitorName, place, country, competitorNation, lenpts = lengthPoints) %>% 
+    mutate(competitionplace = str_sub(place, start=1,end=-6)) %>% 
+    mutate(athome = if_else(country==competitorNation,1,0))
+
+# Model 5: Length model
+model5b <- felm(data=home16, lenpts ~ athome + competitionplace | competitorName)
+summary(model5b)
+
+# Model 6: Style points model
+model6b <- felm(data=scores16, judge ~ athome + competitionplace + judge_name + speed + length| competitorName)
+summary(model6b)
+
 ```
@@
 The experts may completely or partially decide the outcome of competitions.
-Expert judgement is highly valued, and being a judge in sports competitions carries major prestige.
 
 A prerogative of the system is that experts judge performances in an unbiased and neutral manner.
-However, there is a large potential for biases in such judgements (Meyer and Booker 1991).
+However, there is a large potential for biases in such judgements [@meyer_selecting_1991].
 Experts may, for example, overestimate performances by competitors from their own country, overvalue certain stylistic elements, or unconsciously adjust evaluations in response to spectator expectations.
@@
 2008), and in assessing job applicants (cf.
-Fasang 2006; Bursell 2012).
+@fasang_recruitment_2006; @bursell_name_2012).
 Understanding the extent of and processes underlying such biases is important for any attempts to remove their effects.
@@
 Nationalistic bias refers to biases in which the judges (positively) evaluate contenders from their own country, whereas home-field advantage bias arises if judges favour contenders from the country in which the contest is held.
-To answer our questions, we use data from the 2006-07 and 2007-08 seasons of the Federation International de Ski (FIS) World Cup series of ski jumping contests.
+To answer our questions, we use data from two periods, the 2006-2008 seasons and the 2015-2016 season, of the Federation International de Ski (FIS) World Cup series of ski jumping contests.
 
 We contribute to the literature on nationalistic and home field advantage biases in several ways.
-First, by analysing both nationalistic and bias home field advantages on the same data, we can compare different aspects and sources of such bias.
-Second, by separately analysing the social and physical ramifications of home advantages, we can provide additional evidence on the nature of home field advantages.
-Third, by comparing our results to those obtained with similar data but from the markedly different context of the Olympic games, a highest-stakes-contest [@zitzewitz_nationalism_2006], we can shed light on the potential for biased evaluations in different institutional and incentive contexts.
+First, we analyse nationalistic bias and assess to what extent such bias can be explained by unobserved characteristics of the performance.
+Second, we consider variations in such bias, and discuss this variation in light of sociological theory of evaluation and valuation.
+Third, by using our data from two time periods and comparing our results to those obtained with similar data but from the markedly different context of the Olympic games, a highest-stakes-contest [@zitzewitz_nationalism_2006], we can shed light on the role of different institutional and incentive contexts and changes over time in the magnitude of bias.
+Fourth, we study the social and physical ramifications of home advantages.
 
-# A REVIEW OF THE LITERATURE ON SUBJECTIVE BIASES IN SPORTS
+# THEORY AND RELEVANT LITERATURE
 
-A relatively large literature explores the existence and magnitude of subjective biases in sports evaluations.
+There is an extant scholarly literature that explores the existence and magnitude of subjective biases in sports evaluations.
 These biases may affect results or other outcomes in favor of what is in principle irrelevant aspects of competitors or other actors in the sport.
@@
 
-## Nationalistic biases
-
 Nationalistic or cultural bias, where judges favour athletes from their home country or culturally similar ones in a systematic way would be most evident in certain nationalistic competitions such as world cups or Olympics.
@@
 
-Vote trading in the 2002 Olympics led to a revamp of the scoring system for figure skating after a French judge admitted to being pressured to vote in favour of the Russian pair over the Canadian pair in the pairs’ competition final – effectively handing them the gold medal.
+Vote trading in the 2002 Olympics led to a revamp of the scoring system for figure skating after a French judge admitted to being pressured to vote in favour of the Russian pair over the Canadian pair in the pairs’ competition final.
+This effectively handed the Russian pair the gold medal.
 Allegedly this was part of a deal with the Russian team, where they would reciprocate in a later competition in favour of the French.
@@
 
-## Home field advantages
+## Theorizing nationalistic biases in ski jumping
 
-Home field advantage is another type of bias.
+Our reading of the empirical literature on nationalistic and other evaluative biases suggests there is evidence for such biases in a variety of sports.
+Given its existence, what are the root causes of such bias and how do they emerge?
+Theoretically, we can concieve of bias in at least three different ways.
+We denote these three mechanisms the social psychological mechanism, the cultural legitimacy mechanism and the differential professionalization mechanism, and discuss each of them in turn below.
+
+### The social psychological mechanism: Implicit or explicit nationalist bias
+
+The first social psychological mechanism is the simple folk psychology theory of  judges' nationalistic impulses at the subconscious or even conscious level.
+They prefer their own to win, and give higher scores to those athletes.
+Importantly, this theoretical mechanism does not by itself predict variation in the magnitude of biases across countries.
+It could in princple be combined with other theoretical ideas to arrive at predictions about any such variation.
+
+### Theory of evaluation work, as applied to nationalist biases in ski jumping
+
+The second way of theorizing bias is to consider the role of judges and their efforts as evaluation work.
+Evaluative work and the justification for evaluations as a sociological study has been profoundly influences by the works of Boltanski \& Thevenon [@bolstanski_justification_1991].
+A central tenet in their sociology of evaluations is the room for evaluation.
+
+The room for evaluation decides what is relevant and good in a field.
+The evaluative room may be wide, and allow for many different evaluative statements, or narrow, and restrict evaluative statements.
+The main point is that the evaluative room allows for criticisms by field actors that in turn may affect the structure of the room for evaluation.
+Criticisms within the room for evaluation are recognized as forces of change in organizations and society at large.
+Actors cannot expect to have any criticism heard, and it must be recognized by multiple actors before it is heard.
+If enough actors repeat and (re)formulate  a criticism, it will potentially lead to change and a new room for evaluation.
+
+Applied to ski jumping as a field, judges' practices for scoring represents the room for evaluation.
+Within this room there are likely coexisting practices that may be correlated with judges' nationalities, giving rise to national variations in judge scores.
+Some countries, for example, Norway and Finland, enjoy long ski jumping traditions, whereas neighbouring Sweden, an otherwise similar country, does not.
+Most likely the long traditions also imply that judges from these countries to a larger degree than Swedish judges will define the room for evalution.
+This became salient when Jan Boklöv, a Swede, pioneered the V-style in ski jumping.
+His style was not comme-il-faut among judges and competitors in ski jumping at the time, and thus heavily penalized by judges.
+Eventually he won competitions, due to the technical advantages of the V-style, and other actors started to mimic the style [@muller_performance_2008].
+These changes then led to a change in the room for evaluations, where V-style was accepted and it is currently completely dominant in the field.
+
+Mere national variation in judge scoring practices does not yield nationalistic biases in individual scores.
+However, when we also consider the fact that competitors are trained and socialized in their respective national ski jumping associations and competitions.
+This implies that young competitors share the more experienced judges' evaluative standards, and in turn this leads to nationalistic biases in judge scoring.
+
+### Professionalisation of sports and the evolution of evaluation
+
+The third source of national variation in ski jumping evaluations is related to the process of professionalization in sports.
+An essential ingredient in evaluation are standards.
+In sports, standards are continuously evolving.
+The continued professionalization of sports could would suggest a higher degree of standardization and a narrowing of the room for evaluations.
+Several aspects of contemporary ski jumping suggests that professionalization will limit the magnitude of nationalistic biases.
+
+First, FIS licences judges that takes international, high-level judging assignments.
+Professionalization of ski jumping entails stricted quality control of judge performances and training.
+Currently FIS requires national skiing associations to hold at least one national-level judge seminar every year to contribute to educating new judges, and all judges are required to participate in such seminars at least once every two years.
+
+Second, There is also a system in place for auditing judges.
+Chapter 10 of the FIS Rules documents explains that judges are subject to an evaluation by a FIS Sub-committee working together with a data team, and that their work may result in allotting individual judges with "points" that indicate non-fair scoring practices.
+If judges accumulate many points, they may not be nominated to higher-level competitions or wholly excluded from judging.
+Such sanctioned judges may re-qualify for judging assignments after scoring performances from television broadcasts.
+
+Third, the competition rules also limit the influence of such biases by removing the best and worst of the five scores from the total score.
+This certainly limits the magnitude of any bias, but as previous research have found significant biases [@zitzewitz_nationalist_2006], it does not seem to root it completely out.
+
+The relevance of professionalism for the *national variation* in biases emerges when we combine the idea of evaluative work with the likely different levels of professionalism among the national associations that are responsible for judge training.
+In countries where ski jumping is a relatively important sport, such as Finland and Norway, the bias should be less due to an assumed stronger professionalization.
+
+Summing up, it seems meaningful to probe the existence of nationalist biases in different countries, but also study whether nationalist biased have decreased in magnitude over time, due to the general tendency towards increased  professionalization in sports.
+
+In addition to the social psychological, cultural legitimacy, and differential professionalization mechanisms discussed above, there may well be yet other mechanisms that also contribute to nationalistic biases of this kind, for example desires on part of individual judges to grow the sport's influence in their countries and that try to boost their own competitors placements.
+
+## The extant literature on home field advantages
+
+Home field advantage is another type of potential bias in ski jumping.
 Such advantages, where some competitors benefit from advantages from competing on the home field, may arise in many different ways.
@@
 
-### Hypotheses
+## Hypotheses
 
-We propose several hypotheses on biases in international ski jumping competitions, based on the literature reviewed above.
+We propose several hypotheses on biases in international ski jumping competitions, based on the literature and theory reviewed above.
 On the basis of earlier findings, we propose a first hypothesis that there are nationalistic biases in ski jumping.
@@
 
-Our second hypotheses states that the gross bias is partly explained by controlling for characteristics of the jump and athlete.
+Our second hypothesis states that the gross bias is partly explained by controlling for characteristics of the jump and athlete.
 The empirical implication of is that the magnitude of the bias will attenuate once all unobserved factors related to the jump are controlled.
 
-A third hypothesis is that this bias may be partly due to cultural influences on stylistic elements, and that the strength of the bias is lower in countries with strong traditions for ski jumping and higher in countries where the sport is relatively less important.
+A third hypothesis concerns variation in bias across countries.
+Nationalistic biases may vary across countries due to cultural-professional variations, where the strength of the bias is lower in countries with strong  jumping cultures and higher in countries where the sport is more peripheral in the larger sports field.
 
-For home field advantages, we propose that home field advantages are measurable and impacts athletes performances.
-We further distinguish between the social aspects of competing on the home field and the physical aspects of competing on the home field.
-We hypothesize that the physical home field advantage is less salient than the social home field advantage, and thus less important for performances.
+Our fourth hypothesis states that nationalistic biases in ski jumping evaluations decrease over time, possibly as a results of continued professionalization of the sport.
+
+For home field advantages, we propose an additional hypothesis that home field advantages are measurable and impacts athletes performances.
+Finally, we distinguish between the social aspects of competing on the home field and the physical aspects of competing on the home field, and hypothesize that the physical home field advantage is less salient than the social home field advantage, and thus less important for performances.
 
@@
 
-To test the hypotheses proposed above, we use data from the FIS World Cup competitions in ski jumping from the 2006/07 and 2007/08 seasons.
-The original data were downloaded from the FIS web site, coded, and collated into complete data set for analysis.
-Our data include detailed information on each jump in each competition, including the five individual judge scores from the jump along with the identity and nationality of each judge in each competition.
-We also know where the competition was held.
+To test the hypotheses proposed above, we use data from the FIS World Cup competitions in ski jumping from the 2006-2007 and 2007-2008 seasons as well as from the 2015-2016 seasons.
+The original data for 2006-2007 and 2007-2008 seasons were downloaded from the FIS web site, coded, edited for errors, and collated into complete data set for analysis.
+For the season 2015-2016, we wrote automated web scraping software that identified relevant PDF files on the FIS servers, downloaded these files, and scraped them according to our pre-specified instructions.
+The software used to download and scrape the PDF files are documented in the GitHub repository associated with this article.
 
-Our data are now about a decade old, an age that in many social science areas may raise questions about the value of our analysis.
-We have made efforts to obtain newer and larger-scale data from owners of data on skijumping competitions.
-Our requests made to both Federation International de Ski and the Norwegian Skiing Assocation were rejected for legal reasons.
-However, even if more recent data would be slighly preferable, we are not worried about the analysis' "expiry date".
-The reason is that the phenomena under study, nationalist and home field biases, are neither new nor old, but general in nature and similar phenomena are likely to persist in any kind of evaluation setting.
-Our analysis, we would argue, has relevance far beyond the scope of our data, the skijumping sport, and perhaps even beyond the sporting domain.
+Both data sets include detailed information on each jump in each competition, including the five individual judge scores from the jump along with the identity and nationality of each judge in each competition.
+Characteristics of the jump such as speed and length are included, as well as the points allotted.
+The total points for the jump is the sum of the length score and the stylistic score.
+We also know where and when the competition was held.
+For the later period, additional details on wind, starting gate and point adjustments for such conditions are also available.
 
-We exclude team events from both seasons, leaving us with a data set of 51 competitions and 3,772 individual jumps.
-Table 1 outlines the dimensionality of the data set.
-The total score of the jump is the sum of the length score and the stylistic score.
-In our data set, individual judge evaluations of each jump are available, as well as additional technical data on the length and speed of each contestant’s jumps.
+We exclude team events from both seasons, as team events follow a different structure than individual competitions and also because extracting data from team events are markedly more difficult than it is for individual events.
+Table 1 outlines the dimensionality of the two data sets.
 
@@
 # TABLE 1. Descriptive statistic
-descheading <- c("Statistic","Mean / Number"); numdigs = 2
+descheading <- c("Statistic","2005-2007","2015-2016"); numdigs = 2
 thetable <- as.data.frame(rbind(
- c("No. of competitor nations", length(unique(jumps$competitornation))),
- c("No. of unique skijumping hills", length(unique(jumps$competitionplace))),
- c("No. of unique competitors",length(unique(jumps$competitorname))),
- c("No. of unique judges", length(unique(d$judge_name))),
- c("No. of jumps", nrow(jumps)),
- c("No. of competitions", length(unique(jumps$competition))),
- c("Avg. judge score", as.numeric(round(mean(as.numeric(d$judge)),numdigs))),
- c("Avg. jump length", as.numeric(round(mean(d$length),numdigs))),
- c("Avg. points", as.numeric(round(mean(d$points),numdigs))), deparse.level=0 ))
+ c("","Mean / Count", "Mean / Count"),
+ c("No. of competitor nations", length(unique(jumps$competitornation)), length(unique(season2016$competitorNation))),
+ c("No. of unique skijumping hills", length(unique(jumps$competitionplace)),length(unique(season2016$place))),
+ c("No. of unique competitors",length(unique(jumps$competitorname)), length(unique(season2016$competitorName))),
+ c("No. of unique judges", length(unique(d$judge_name)), length(unique(scores16$judge_name))),
+ c("No. of jumps", nrow(jumps), nrow(season2016)),
+ c("No. of competitions", length(unique(jumps$competition)), length(unique(season2016$componr))),
+ c("Avg. judge score", as.numeric(round(mean(as.numeric(d$judge)),numdigs)), as.numeric(round(mean(as.numeric(scores16$judge)),numdigs))),
+ c("Avg. jump length", as.numeric(round(mean(jumps$length),numdigs)), as.numeric(round(mean(season2016$length),numdigs))),
+ c("Avg. points", as.numeric(round(mean(jumps$points),numdigs)), as.numeric(round(mean(season2016$totalPoints),numdigs))), deparse.level=0))
 colnames(thetable) <- descheading
-pander(thetable)
+pander(thetable, caption = "Descriptive statistics")
 ```
 
-The analysis proceeds in several steps.
-First, we estimate nationalistic biases using a sequence of regression models where more complexity is incorporated for each model.
-Then, we estimate home field advantage.
-Finally, we simulate how important nationalistic biases and home field advantages are in final rankings of the competitions.
+Our analysis proceeds in several steps.
+First, we estimate nationalistic biases using a sequence of regression models where more complexity is incorporated for each model, each model speaking to one of the first three hypoteses.
+We estimate the models separately for the two periods we have data for, as there may be have been change over time in the nature of nationalistic biases, the subject of our fourth hypothesis.
+In a second step, we estimate home field advantages in both jump performance and stylistic evaluation.
 
@@
 
-Following the general approach used in most studies of nationalistic biases (in particular, Emerson et al.
-(2009) and Zitzewitz (2006)), we analyse systematic tendencies in scores by individual judges and by the judges’ nationality.
+Following the general approach used in most studies of nationalistic biases (in particular, @emerson_assessing_2009; @zitzewitz_nationalism_2006), we analyse systematic tendencies in scores by individual judges and by the judges’ nationality.
 We begin with a simple model and add complexity in subsequent steps to approach a stricter test of nationalistic bias.
@@
 \begin{equation}
-s_{ijp} = B∙ \phi(I=J) + \epsilon_{ijp}
+s_{ijp} = B ∙ \phi(I=J) + \epsilon_{ijp}
 \end{equation}
@@
 However, such a model does not take into account relevant differences between nations.
-A Norwegian judge may give higher scores to a fellow countryman’s performance due to a correspondence of their definitions of valuable stylistic elements.
-For example, Norway and Finland enjoy long ski jumping traditions, whereas neighbouring Sweden, an otherwise similar country, does not.
-The importance of tradition became salient when Jan Boklöv, a Swede, pioneered the V-style in ski jumping but was consistently disadvantaged by the judges.
-His style was not comme-il-faut among judges and competitors in ski jumping at the time, and thus heavily penalized by judges.
+A Norwegian judge may give higher scores to a fellow countryman’s performance due to a correspondence of their definitions of valuable stylistic elements, in accordance with national variations in evaluation practices.
 
@@
 In order to observe a home field effect in points, be it length points or the judges’ style points, we have to compare scores on a home field with scores on away fields – leading to the assumption that a jumper is somewhat consistent in performance throughout a season.
-Previously, we used the four other judges as a more objective measure of performance quality.
+Previously, we used the four other judges as a contrast measure of performance quality.
 For home field advantage we have to assume that an athlete maintains a relatively constant quality of jumping throughout the season, using his away jumps as a guide to his general performance level.
@@
 This is because “all else being equal, a ski jumper with the same take-off speed but better form while airborne will travel further” [@zitzewitz_nationalism_2006, 78].
-We therefore add these two variables to the model, where they are represented by the x.
+We therefore add these two variables to the model, where they and their coefficients are represented by the $βx_p$.
 
@@
 
-Our analysis of nationalistic bias proceeds in four steps.
-Table 2 presents the results from the first three estimations.
-The fluctuation in the number of observations is due to jumps with missing values on one or more of the independent variables, which have been deleted from the estimation.
+Our analysis of nationalistic bias proceeds in four steps. Table 2 presents the results from the first three estimations.
 
@@
     single.row=TRUE,
-    column.labels = c("Model 1: Naïve", "Model 2: Within-jump", "Model 3: Within-judge"),
+    column.labels = c("Model 1a: Naïve", "Model 2a: Within-jump", "Model 3a: Within-judge"),
     title="Results from regression models of national bias",
@@
     notes.align="l",
-    notes = c("In models 2 and 3, the intercept is suppressed by the fixed effects estimation procedure."),
+    notes = c("In models 2a and 3a, the intercept is suppressed by the fixed effects estimation procedure."),
+    notes.append=TRUE,
+    df = TRUE, decimal.mark =".",
+    ci=TRUE, ci.level=0.95, ci.separator = "  ",
+    omit.stat=NULL,
+    star.cutoffs = c(0.05,0.01,0.001)
+    )
+```
+\end{landscape}
+
+\begin{landscape}
+```{r tab2b, echo=FALSE, results='asis'}
+stargazer(model1b, model2b, model3b, out.header=FALSE, header=FALSE,
+    single.row=TRUE,
+    column.labels = c("Model 1b: Naïve", "Model 2b: Within-jump", "Model 3b: Within-judge"),
+    title="Results from regression models of national bias for season 2015-2016",
+    covariate.labels = c("National bias"),
+    dep.var.caption  = "", dep.var.labels.include = FALSE,
+    model.numbers = FALSE, model.names = FALSE,
+    add.lines = c(" "),
+    notes.align="l",
+    notes = c("In models 2b and 3b, the intercept is suppressed by the fixed effects estimation procedure."),
     notes.append=TRUE,
@@
     ggtitle("Figure 1. Histogram of judges' leniency")  
-figure1
+#figure1
+
+judge_effectsb <- filter(as.data.frame(getfe(model3b, se = FALSE)),fe=="judge_name")
+figure1b <- ggplot(data=judge_effectsb, aes(x=effect)) + 
+    geom_histogram(bins=15) +
+    xlab("Leniency") + ylab("Number of judges") +
+    theme_minimal() +
+    ggtitle("Figure 1. Histogram of judges' leniency")  
+#figure1b
+
+botheffects <- list("2005-2007" = judge_effects, "2015-2016" = judge_effectsb) %>% 
+    map(function(df) select(df, effect)) %>% 
+    bind_rows(.id="Season")
+figure1c <- ggplot(data=botheffects, aes(x=effect, fill=Season)) + 
+    geom_histogram(bins=20, alpha=0.5) +
+    xlab("Leniency") + ylab("Number of judges") +
+    theme_minimal() +
+    theme(legend.position = "bottom") + 
+    ggtitle("Figure 1. Histogram of judges' leniency")  
+figure1c
 ```
@@
     mutate(country = str_sub(term, start=-3, end=-1)) %>% 
-    filter(!(country %in% c("USA","SVK","SWE","ITA"))) %>%
+#    filter(!(country %in% c("USA","SVK","SWE","ITA"))) %>%
     mutate(ocountry = factor(country, levels = .$country[rev(order(.$estimate))])) %>% 
@@
     geom_point(size=3) + 
-    geom_pointrange(aes(x=country, y=estimate, ymin = lowci, ymax=hici), fatten = 1.5) +
+    geom_pointrange(aes(x=country, y=estimate, ymin = lowci, ymax=hici), fatten = 1.2) +
     #color="grey") +
@@
     ggtitle("Figure 2. Nation-specific estimates of judges' bias")
-figure2
+#figure2
+
+nationbiasb <- broom::tidy(model4b) %>% 
+    mutate(country = str_sub(term, start=-3, end=-1)) %>% 
+    mutate(ocountry = factor(country, levels = .$country[rev(order(.$estimate))])) %>% 
+    mutate(lowci = estimate-1.96*std.error, hici=estimate+1.96*std.error) %>% 
+    arrange(estimate)
+
+figure2b <- ggplot(data = nationbiasb,mapping = aes(x = ocountry, y = estimate)) +
+    geom_point(size=3) +  
+    geom_pointrange(aes(x=country, y=estimate, ymin = lowci, ymax=hici), fatten = 1.2) +
+    #color="grey") +
+    #geom_segment(aes(x=country, y=estimate, xend=country, yend=0)) +
+    xlab("Nation") + ylab("Strength of bias") +
+    theme_minimal() +
+    ggtitle("Figure 2. Nation-specific estimates of judges' bias")
+#figure2b
+
+bothbias <- list("2005-2007" = nationbias, "2015-2016" = nationbiasb) %>% 
+    map(function(df) select(df, country, estimate, lowci, hici)) %>% 
+    bind_rows(.id="Season") %>% 
+    arrange(Season, estimate)
+figure2c <- ggplot(data = bothbias,mapping = aes(x = forcats::fct_rev(forcats::fct_reorder(country, estimate)), y = estimate, color=Season)) +
+    geom_point(size=3) +  
+    geom_pointrange(aes(x=country, y=estimate, ymin = lowci, ymax=hici), fatten = 1.2) +
+    xlab("Nation") + ylab("Strength of bias") +
+    theme_minimal() +
+    theme(legend.position = "bottom") + 
+    ggtitle("Figure 2. Nation-specific estimates of judges' bias")
+figure2c
 ```
@@
 The biases were generally small, and do not seem to pose a major threat to the validity of results from skijumping competitions.
-However, in league or series-type competition tournaments, the cumulative effects of natioanlistic bias may be important.
-Future research should emphasize this problem.
+However, in league or series-type competition tournaments, the cumulative effects of nationalistic bias may be important.
+An important takeaway from this study is that the system of regulations in
 
 Biases in subjective evaluations of judges and referees may carry important implications for sports results, as exemplified by findings of potential effects on final rankings of athletes or on betting markets [@larsen_racial_2008, @price_racial_2010].
-Given the historically political nature of sports—as exemplified in extreme cases by the boycotts of the United States and the Soviet Union of the Moscow and Los Angeles Olympics, respectively, the contested nature of the 1972 chess championships between Fisher and Spassky, and the many cases of violence between sports fans (and, at the most extreme end the alleged “Football War” between Honduras and El Salvador in 1969)—subjective biases in sports can even carry political implications.
-Knowledge of the sources of evaluative biases in competitive sports can help us better understand biases and discriminatory practices in other fields, such as student and employee recruitment and the allocation of research grants.
+
+Knowledge of the sources of evaluative biases in competitive sports can potentially help us better understand biases and discriminatory practices in other fields, such as student and employee recruitment and the allocation of research grants.
 Our findings of nationalistic and home-field biases have provided evidence that, in particular when assessed together with other related findings, add to our understanding on bias-generating practices.

### 2019-07-25 scrape/download-fis-files.R

@@ -1,0 +1,37 @@
+##
+## Download all the results files from FIS
+## URLs and other info is read from file.
+##
+library(tidyverse)
+library(xml2)
+library(httr)
+library(rvest)
+library(here)
+
+### Read in scraped URLs 
+allseasons <- data.table::fread(here("scrape/scraped-url-data.csv")) %>% 
+    mutate(fileid = str_c(season,"-",eventid,"-",raceid,"-",codex,"-",filetype,".pdf"))
+summary(allseasons)
+
+targets <- allseasons %>% 
+    filter(filetype=="RL1")
+
+### Create directory structure
+###
+if (!dir.exists(here("scrape/pdfs"))) dir.create(here("scrape/pdfs"))
+
+downloadPDFsafe <- function(desc) {
+    destpath <- here(str_c("scrape/pdfs/",desc$fileid))
+    if (file.exists(destpath) && file.size(destpath)>0) {
+        warning(str_c("File ", desc$fileid, " already exists. Skipping!"))
+    } else {
+        download.file(url = desc$pdf_url, destfile = destpath, timeout=5)
+        Sys.sleep(5)
+    }
+}
+
+targets %>% 
+    select(pdf_url, fileid) %>% 
+    mutate(n = row_number()) %>% 
+    split(.,.$n) %>% 
+    map(downloadPDFsafe)

### 2019-07-25 scrape/obtain-pdf-urls.R

@@ -1,0 +1,130 @@
+##
+## Download all the results files from FIS
+##
+library(tidyverse)
+library(xml2)
+library(httr)
+library(rvest)
+library(here)
+
+###
+### Function to generate list of a season's WC competitions
+###
+getSeason <- function(yr) {
+    return(str_c("https://www.fis-ski.com/DB/general/event-details.html?eventselection=&place=&sectorcode=JP&seasoncode=", as.character(yr), "&categorycode=WC&disciplinecode=&gendercode=&racedate=&racecodex=&nationcode=&seasonmonth=X-", as.character(yr), "&saveselection=-1&seasonselection="))
+}
+
+###
+### Function to scrape URLs for different events for one season, based on season list URL
+###
+scraplinks <- function(url){
+    # Create an html document from the url
+    webpage <- xml2::read_html(url)
+    # Extract the URLs
+    url_ <- webpage %>%
+        rvest::html_nodes("a") %>%
+        rvest::html_attr("href")
+    # Extract the link text
+    link_ <- webpage %>%
+        rvest::html_nodes("a") %>%
+        rvest::html_text()
+    return(tibble(link = link_, url = url_))
+}
+
+###
+### Function to filter links scraped by scraplinks to get urls to season events
+###
+filterUrls <- function(links, ssn) {
+    theUrls <- links %>% 
+        filter(str_detect(url, "eventid=")) %>% 
+        filter(str_detect(link, c(as.character(as.numeric(ssn-1)),as.character(ssn)))) %>% 
+        select(url)
+    return(theUrls)
+}
+
+###
+### Function that produces a list of urls of season events from a season year argument
+###
+seasonEvents <- function(season) {
+    getSeason(season) %>% 
+        scraplinks %>% 
+        filterUrls(.,ssn = season)
+}
+
+###
+### Function that filters all URLs for one event into unique links for the races in that event
+###
+filterRaces <- function(links) {
+    theUrls <- links %>% 
+        filter(str_detect(url, "raceid=")) %>%
+        filter(str_detect(url, "results.html")) %>% 
+        select(url)
+    return(theUrls)
+}
+
+###
+### MAIN FUNCTION FOR GETTING ONE SEASON OF DATA
+###
+getURLdata <- function(season) {
+
+    ###
+    ### List of events generated from season summary calendar page
+    ###
[…]

### 2019-07-25 scrape/scrape-2016-format.R

@@ -1,0 +1,183 @@
+###
+### Program for parsing 2016 season format PDFs
+###
+library(tabulizer)
+library(pdftools)
+library(magick)
+library(tidyverse)
+library(here)
+
+
+### 
+### Functions to obtain header information
+###
+get_compo_info <- function(thefile) {
+    
+    competitionName_xy <- c(top=23,left=140,bottom=45,right=510)
+    competitionDate_xy <- c(top=48,left=400,bottom=68,right=500)
+    competitionPlace_xy <- c(top=63,left=140,bottom=80,right=384)
+    competitionType_xy <- c(top=79, left=143,bottom=95,right=305)
+    
+    componame <- tabulizer::extract_tables(file = here("scrape/pdfs/",thefile), pages = 1, guess=F, area = list(competitionName_xy))
+    compodate <- tabulizer::extract_tables(file = here("scrape/pdfs/",thefile), pages = 1, guess=F, area = list(competitionDate_xy))
+    compoplace <- tabulizer::extract_tables(file = here("scrape/pdfs/",thefile), pages = 1, guess=F, area = list(competitionPlace_xy))
+    compotype <- tabulizer::extract_tables(file = here("scrape/pdfs/",thefile), pages = 1, guess=F, area = list(competitionType_xy))
+    
+    compoinfo <- data.frame(series = componame[[1]], place = compoplace[[1]], date = compodate[[1]], ctype = compotype[[1]])
+    return(compoinfo)
+}
+
+###
+### Judge information
+###
+get_judge_info <- function(thefile) {
+    
+    judges_xy <- c(top = 130, left = 236,bottom = 190,right = 375)
+    blurb <- tabulizer::extract_tables(file = here("scrape/pdfs/",thefile), pages = 1, guess=F, area = list(judges_xy))
+    
+    judges <- blurb %>%
+        as.data.frame %>% 
+        setNames(nm=c("judgeNo","namenat")) %>% 
+        mutate(judgeName = str_sub(namenat, start=1, end=-7)) %>% 
+        mutate(judgeNation = str_sub(namenat, start=-4,end=-2)) %>%
+        select(-namenat)
+    judges_wide <- judges %>% 
+        gather(key=var, value=val, -judgeNo) %>% 
+        mutate(var = str_c(var,judgeNo)) %>% 
+        select(-judgeNo) %>% 
+        spread(key=var, value=val)
+    
+    return(judges_wide)
+}
+
+###
+### MAIN CODE SEGMENT:
[…]

### 2019-07-25 scrape/scraping_attempt.R

@@ -1,0 +1,99 @@
+library(tabulizer)
+library(pdftools)
+library(magick)
+library(tidyverse)
+library(here)
+
+
+
+### Read list of links scraped and downloaded
+pdfs <- list.files(here("scrape/pdfs"), pattern="*.pdf") 
+thefile <- sample(grep(pdfs, pattern="2016", value=T),1)
+
+# # Convert first page to png
+# dir.create(here("temp"))
+# pdf_convert(here("newer-data/",thefile), pages=1, dpi=300, filenames = here("temp/firstpage.png"))
+# # Cut out judge names
+# firstpage <- image_read(here("temp/firstpage.png"))
+#     judge_table <- image_crop(firstpage, "400x200+1000+560")
+# image_write(judge_table, path = "temp/judgetable.png")
+# system("convert temp/judgetable.png temp/judgetable.pdf")
+
+
+### Pull out information about the judges from the PDFs
+# This area is for the newest format
+judges_xy <- c(top = 130, left = 236,bottom = 190,right = 375)
+blurb <- tabulizer::extract_tables(file = here("scrape/pdfs/",thefile), pages = 1, guess=F, area = list(judges_xy))
+blurb
+
+judges <- blurb %>%
+    as.data.frame %>% 
+    setNames(nm=c("judgeNo","namenat")) %>% 
+    mutate(judgeName = str_sub(namenat, start=1, end=-7)) %>% 
+    mutate(judgeNation = str_sub(namenat, start=-4,end=-2)) %>%
+    select(-namenat)
+judges_wide <- judges %>% 
+    gather(key=var, value=val, -judgeNo) %>% 
+    mutate(var = str_c(var,judgeNo)) %>% 
+    select(-judgeNo) %>% 
+    spread(key=var, value=val)
+judges
+judges_wide
+
+### Get first page results
+first_results_xy <- c(top = 231, left = 28,bottom = 677,right = 567)
+blurb <- tabulizer::extract_tables(file = here("scrape/pdfs",thefile), pages = 1, guess=F, area = list(first_results_xy))
+
+thenames<-c("placement","number","competitor","competitorNation","speed","length","lengthPoints","judgeA","judgeB","judgeC","judgeD","judgeE","stylePoints","startGate","gatePoints","windSpeed","windPoints","totalPoints")
[…]

### 2019-08-12 revised-manuscript/response_to_reviewers.Rmd
Revising ongoing; now new analysis

@@
 
-We have expanded our discussion of home field advantages somewhat in this version, and make amongst other things the discintion you point to in your comment. 
-TODO: OR SHOULD WE JUST DROP HOME FIELD?
+In order to provide a more focused and coherent manuscript, we have removed the home field advantage aspect from this paper.
+Our decision was based on the fact that we expanded the nationalist bias analysis, and this expanded the scope of the work outside of a single article.
+We may proceed with a deeper analysis of home advantages in a subsequent paper.
+
+If the Editor and the reviewers found much promise in the home field advantage component, we are of course willing to put it back in.
 

### 2019-08-12 revised-manuscript/manuscript.Rmd
Revising ongoing; now new analysis

@@
 ---
-title: "NATIONALIST AND HOME FIELD BIAS IN SPORT PERFORMANCE EVALUATIONS"
+title: "NATIONALIST BIAS IN SPORT PERFORMANCE EVALUATIONS"
 subtitle: "An example from the ski jumping World Cup"
@@
 date: "14 september 2017"
-keywords: nationalist bias, home field bias, skijumping, judges
+keywords: nationalist bias, skijumping, judges, fixed effects
 abstract: "High stakes and prizes in international sports competitions necessitate efficiently identifying legitimate winners.
-Illegitimate biases of non-independent judges and home field advantages represent two major threats to the legitimacy of international competitions.
+Illegitimate biases of non-independent judges represent a major threat to the legitimacy of international competitions.
 Judges’ nationalities represent one source of bias, as they may prefer athletes of their own nation.
@@
 We draw on data from the FIS World Cup competitions in the 2006-2008 and 2015-2016 seasons, and employ several fixed-effects based regression designs.
-Our results reaffirm suspicions of national bias and home field advantage in major ski jumping competitions, but also show their magnitude is too little to be of major relevance to competition outcomes.
-There is a trend towards smaller nationalist biases  "
+Our results reaffirm suspicions of national bias in major ski jumping competitions, but also show their magnitude is too little to be of major relevance to competition outcomes.
+There is a trend towards smaller nationalist biases over time, which suggests that increasing professionalization of sports also may limit the influence of subjective and nationalist impulses."
 thanks: "The authors are grateful for financial support from the Småforsk programme at the University of Oslo, and to Marte Knutsson for research assistance in the early phase of establishing the data set.
@@
 
+### SEASON 2005-06 DATA FROM EARLY ANALYSIS
 # Read in data
@@
 
-# Correct judge names
+### Correct judge name spellings
 pairs <- matrix(c("YERLY James","YERRLY James","ASPHAUG Kolbjoern","ASPHAUG Kolbjoer","DALLE AVE Sandro","DALLE Ave Sandro","EGLOFF Ernst","EGLOFF Emst","KARJALAINEN Pirjo","KARJALAINEN Pirvo","KUUSINEN Eero","KUUSINE Eero","LOENG Geir","LOENG Geir S.","LORENZ Michael","LORENTZ Michael","NYMAN Tom","NYMAN Torn","SCHMIEDER Jorg","SCHMIEDER Joerg","SCHMIEDERER Konrad","SCHIMIEDERER Konrad"), ncol=2, byrow=TRUE)
@@
 
-# Various data preparation
+### Various data preparation
 d <- j %>% 
@@
   mutate(athome = ifelse(competitornation==country,1,0)) %>%
-  mutate(competitorname = as.factor(competitorname)) %>%
-  mutate(competitionplace = as.factor(competitionplace)) %>%
+  mutate(competitorname = tolower(competitorname)) %>%
   mutate(athseason = as.factor(str_c(competitorname,as.character(season)))) %>%
+  mutate(judge = as.numeric(judge)) %>% 
   # Drop unneeded variables
@@
 #
-# ESTIMATIONS: NATIONAL BIAS  MODELS
-#
-# Model 1, Naïve model of national bias
-model1 <- lm(data = d,judge ~ natbias)
-summary(model1)
-
-# Model 2: introducing jump fixed effects
-model2 <- felm(data = d, judge ~ natbias | jumpid)
-summary(model2)
-
-# Model 3: Introducing judge fixed effects
-model3 <- felm(data = d, judge ~ natbias  | jumpid + judge_name)
-summary(model3)
-
-# Model 4: Does national bias vary among nations?
-model4 <- felm(data = d, judge ~ relevel(countrybias,ref = "NON") | jumpid + judge_name)
-summary(model4)
-
-# 
-# * HOME FIELD BIAS MODELS
-# 
-# Model 5: Length model
-e <- filter(d, judgeno==1)
-model5 <- felm(data=e, lenpts ~ athome + competitionplace | athseason)
-summary(model5)
-
-# Model 6: Style points model
-model6 <- felm(data=d, judge ~ athome + competitionplace + judge_name + speed + length| athseason)
-summary(model6)
-
-#
 # Season 2015-2016 analysis
 #
-season2016 <- data.table::fread(here("scrape/scraped-data-season-2016.csv"))
-
+season2016 <- data.table::fread(here("scrape/scraped-data-season-2016.csv"), stringsAsFactors = F)
 scores16 <- season2016 %>% 
-    select(contains("judge"), competitorNation, country, place, speed, length, competitorName, gender) %>%
+    select(contains("judge"), competitornation = competitorNation, country, place, speed, length, competitorname = competitorName, gender) %>%
     mutate(jumpid = row_number()) %>%
@@
     ungroup %>% 
-    mutate(natbias = if_else(competitorNation==judge_nation,1,0)) %>%
+    mutate(natbias = if_else(competitornation==judge_nation,1,0)) %>%
     mutate_at(c("judge"), as.numeric) %>% 
     mutate(countrybias = factor(ifelse(natbias==1,as.character(judge_nation),"NON"))) %>% 
-    mutate(competitionplace = str_sub(place, start=1,end=-6)) %>% 
-    mutate(athome = if_else(country==competitorNation,1,0))
+    mutate(competitionplace = str_sub(place, start=1,end=-7)) %>% 
+    mutate(athome = if_else(country==competitornation,1,0)) %>% 
+    mutate(competitorname = tolower(competitorname))
 
-# NATIONAL BIAS MODELS WITH 2016 DATA
+### Combine seasons
+###
+###
+scores06 <- setNames(d, tolower(names(d))) 
+scores16 <- setNames(scores16, tolower(names(scores16)))
+
+all_scores <- bind_rows(scores06,scores16, .id="whichseason") %>% 
+    select(whichseason, jumpid, judge, judge_name, judge_nation, natbias, countrybias, competitionplace, competitorname) %>% 
+    mutate(theseason = if_else(whichseason==1, "2005-06","2015-16")) %>% 
+    mutate(seasondummy = if_else(whichseason==1,0,1))
+
+# ESTIMATIONS: NATIONAL BIAS MODELS
 #
-### Model 1b: Naïve model
-model1b <- lm(data=scores16, judge ~ natbias)
-summary(model1b)
+# Model 1, Naïve model of national bias
+model1 <- lm(data = all_scores,judge ~ natbias)
+summary(model1)
 
-### Model 2b: Introducing jump fixed effects
-model2b <- felm(data = scores16, judge ~ natbias | jumpid)
+# Model 2: introducing jump fixed effects
+model2 <- felm(data = all_scores, judge ~ natbias | jumpid)
+summary(model2)
+
+# Model 2 with judge fixed effects for looking at leniency
+model2b <- felm(data = all_scores, judge ~ natbias | jumpid + judge_name)
 summary(model2b)
 
-### Model 3b: Introducing judge fixed effects
-model3b <- felm(data = scores16, judge ~ natbias  | jumpid + judge_name)
-summary(model3b)
+# Model 3: Does national bias vary among nations?
+all_scores$countrybias_ord <- relevel(as.factor(all_scores$countrybias), ref="NON")
+model3 <- felm(data = all_scores, judge ~ relevel(countrybias_ord,ref = "NON") | jumpid)
+summary(model3)
 
-# Model 4b: Does national bias vary among nations?
-model4b <- felm(data = scores16, judge ~ relevel(countrybias,ref = "NON") | jumpid + judge_name)
-summary(model4b)
+# Model 4a: Estimating changes over time
+model4a <- felm(data = all_scores, judge ~ natbias + natbias*seasondummy + seasondummy| jumpid)
+summary(model4a)
 
-# HOME FIELD BIAS MODELS WITH 2016 DATA
-# 
-home16 <- season2016 %>% 
-    select(competitorName, place, country, competitorNation, lenpts = lengthPoints) %>% 
-    mutate(competitionplace = str_sub(place, start=1,end=-6)) %>% 
-    mutate(athome = if_else(country==competitorNation,1,0))
+# Model 4b: Estimating changes in leniency and national variance
 
-# Model 5: Length model
-model5b <- felm(data=home16, lenpts ~ athome + competitionplace | competitorName)
-summary(model5b)
-
-# Model 6: Style points model
-model6b <- felm(data=scores16, judge ~ athome + competitionplace + judge_name + speed + length| competitorName)
-summary(model6b)
+## Alternative model 4 with individual-level bias estimates. Not included in the final paper.
+model4_indjudges <- felm(data=all_scores, formula=judge ~  natbias*judge_name | jumpid) 
+model4_indjudges %>% 
+    tidy %>% 
+    mutate(natbias = if_else(str_detect(term,"natbias"),"interaction","main")) %>% 
+    mutate(judge_name = str_sub(term, start=1+((natbias=="interaction")*8))) %>% 
+    mutate(judge_name = if_else(judge_name=="", "overall",judge_name)) %>% 
+    select(natbias, judge_name, estimate) %>% 
+    group_by(judge_name) %>% 
+    spread(key=natbias, estimate) %>% 
+    rename(bias = interaction, leniency = main) %>% 
+    ungroup %>% 
+    mutate(judge_name = str_sub(judge_name,start=11)) %>% 
+    head
 
@@
 
-We contribute to the literature on nationalistic and home field advantage biases in several ways.
-First, we analyse nationalistic bias and assess to what extent such bias can be explained by unobserved characteristics of the performance.
-Second, we consider variations in such bias, and discuss this variation in light of sociological theory of evaluation and valuation.
+We contribute to the literature on nationalistic biases in several ways.
+First, we assess to what extent nationalistic bias can be explained by unobserved characteristics of the performance.
+Second, we consider national variations in such bias, and discuss this variation in light of sociological theory of evaluation and valuation.
 Third, by using our data from two time periods and comparing our results to those obtained with similar data but from the markedly different context of the Olympic games, a highest-stakes-contest [@zitzewitz_nationalism_2006], we can shed light on the role of different institutional and incentive contexts and changes over time in the magnitude of bias.
-Fourth, we study the social and physical ramifications of home advantages.
 
@@
 
-Our reading of the empirical literature on nationalistic and other evaluative biases suggests there is evidence for such biases in a variety of sports.
+Our reading of the empirical literature on nationalistic and other evaluative biases concludes that there is evidence for such biases in a variety of sports.
 Given its existence, what are the root causes of such bias and how do they emerge?
-Theoretically, we can concieve of bias in at least three different ways.
+Theoretically, we can concieve of bias resulting from at least three different mechanisms.
 We denote these three mechanisms the social psychological mechanism, the cultural legitimacy mechanism and the differential professionalization mechanism, and discuss each of them in turn below.
@@
 The second way of theorizing bias is to consider the role of judges and their efforts as evaluation work.
-Evaluative work and the justification for evaluations as a sociological study has been profoundly influences by the works of Boltanski \& Thevenon [@bolstanski_justification_1991].
+Evaluative work and the justification for evaluations as a sociological study has been profoundly influences by the works of Boltanski \& Thevenon [@boltanski_justification_1991].
 A central tenet in their sociology of evaluations is the room for evaluation.
@@
 Third, the competition rules also limit the influence of such biases by removing the best and worst of the five scores from the total score.
-This certainly limits the magnitude of any bias, but as previous research have found significant biases [@zitzewitz_nationalist_2006], it does not seem to root it completely out.
+This certainly limits the magnitude of any bias, but as previous research have found significant biases [@zitzewitz_nationalism_2006], it does not seem to root it completely out.
 
@@
 
-Summing up, it seems meaningful to probe the existence of nationalist biases in different countries, but also study whether nationalist biased have decreased in magnitude over time, due to the general tendency towards increased  professionalization in sports.
+Summing up, it seems meaningful to reexamine the existence of nationalist biases, study how these biases vary across the countries from which evaluators are drawn, and study whether nationalist biased have decreased in magnitude over time, due to, for example, the general tendency towards increased professionalization in sports.
 
 In addition to the social psychological, cultural legitimacy, and differential professionalization mechanisms discussed above, there may well be yet other mechanisms that also contribute to nationalistic biases of this kind, for example desires on part of individual judges to grow the sport's influence in their countries and that try to boost their own competitors placements.
-
-## The extant literature on home field advantages
-
-Home field advantage is another type of potential bias in ski jumping.
-Such advantages, where some competitors benefit from advantages from competing on the home field, may arise in many different ways.
-Several of the mechanisms that may produce home field advantages are pscyhological sources of superior performances, while others involve judges overrating home field competitors.
-This point illustrates that the mere presence of a biases is not necessarily illegitimate.
-In order to improve scoring and fairness in sports, one must understand the mechanisms behind them.
-
-Some research into the home-field advantage has attempted to move our understanding of the causal mechanisms behind  the advantages.
-Carmichael and Thomas [-@carmichael_homefield_2005] outline the four most common explanations of this phenomena using soccer as an example.
-The first is a familiarity explanation – the home team is more familiar with the field, the stadium, the locker room, and so forth, which puts the players more at ease in the situation, allowing them to concentrate on performing at their best.
-Second, traveling to another field could result in fatigue, invoke various stressors, and cause disruptions in pre-competition routines for the away team.
-All these factors may lead to inferior performances by the away team, relative to their likely performance at the home field.
-Third, specific rules may favour the home team, giving them the upper hand.
-Last, there could be a crowd effect.
-The size of the crowd at the home field might improve individual or team performance.
-It might also make the judge to be more lenient for the home team, through psychological mechanisms.
-
-Page and Page [-@page_evidence_2010] attributes some of the crowd effect to crowd noise, noting that the judge is under both time pressure (making a correct decision in a short time) and social pressure (from the crowd).
-A study by Nevill et al.
-[-@nevill_influence_2002] showed that crowd noise was related to the home-field advantage.
-Setting up an experiment with one group of judges watching a football match with sound, and another group without, they found that the group watching with sound called 15.5% fewer fouls against the home team than the group who watched in silence.
-This could also indicate that judges use crowd noise as a guideline in their judging work.
-The crowd effect may also take another direction, as Anders and Rotthoff [-@anders_homefield_2014] find that football clubs with notoriously violent fans have a larger home-field advantage than peaceful ones.
-Here, the theory is that the prospect of fans-to-judge violence causes a more lenient judge.
-
-Research on the home-field advantage has also been done in various winter sports.
-Balmer et al.
-[-@balmer_home_2001] analysed home-field advantages the winter Olympics from 1908 to 1998, and found some evidence for it in figure skating, freestyle skiing, ski jumping, alpine skiing and short track speed skating.
-The authors examined the relevance of three of the explanations listed above (familiarity with local conditions, time zones and crowd effect), and found evidence for home condition familiarity.
-Conditional on the time zones travelled, travel time as such did not matter.
-Most evidence, however, was found for the crowd effect theory, as sports that were judged only subjectively had a bigger home-field advantage than sports in which assessments were based on both subjective and objective criteria.
 
@@
 A third hypothesis concerns variation in bias across countries.
-Nationalistic biases may vary across countries due to cultural-professional variations, where the strength of the bias is lower in countries with strong  jumping cultures and higher in countries where the sport is more peripheral in the larger sports field.
+Nationalistic biases may vary across countries due to cultural, historical and professional ideosyncracies at the national level.
+The strength of the bias is assumed to be lower in countries without strong ski jumping cultures and higher in countries where the sport is more peripheral in the larger sports field.
 
-Our fourth hypothesis states that nationalistic biases in ski jumping evaluations decrease over time, possibly as a results of continued professionalization of the sport.
-
-For home field advantages, we propose an additional hypothesis that home field advantages are measurable and impacts athletes performances.
-Finally, we distinguish between the social aspects of competing on the home field and the physical aspects of competing on the home field, and hypothesize that the physical home field advantage is less salient than the social home field advantage, and thus less important for performances.
+Our fourth hypothesis states that nationalistic bias in ski jumping evaluations decrease over time, possibly as a result of the continued professionalization of the sport.
 
@@
 The original data for 2006-2007 and 2007-2008 seasons were downloaded from the FIS web site, coded, edited for errors, and collated into complete data set for analysis.
-For the season 2015-2016, we wrote automated web scraping software that identified relevant PDF files on the FIS servers, downloaded these files, and scraped them according to our pre-specified instructions.
-The software used to download and scrape the PDF files are documented in the GitHub repository associated with this article.
+For the season 2015-2016, we wrote automated web scraping software that identified relevant PDF files on the FIS servers, downloaded these files, and scraped them for results according to our pre-specified instructions.
+In accordance with norms of open social science, the software used to download and scrape the PDF files are documented in the GitHub repository associated with this article.[^1: See https://github.com/torkildl/skijumping for data and software.]
 
@@
 
-## Estimation of nationalistic biases
+## Estimation of nationalistic bias using fixed effects models
 
 Following the general approach used in most studies of nationalistic biases (in particular, @emerson_assessing_2009; @zitzewitz_nationalism_2006), we analyse systematic tendencies in scores by individual judges and by the judges’ nationality.
-We begin with a simple model and add complexity in subsequent steps to approach a stricter test of nationalistic bias.
-In the first, simple, model, a dummy variable $\phi(I=J)$ indicates whether the judge comes from the same country (J) as the athlete (I) or not.
-The model can be written as
+We begin with a simple model and add complexity in subsequent steps to test our different hypotheses.
+
+In the first, simple, model, a dummy variable $\phi(I=J)$ indicates whether the judge comes from the same country *j* as the athlete *i* or not. The model can be written as
 
@@
 The $B$ coefficient in this model estimates the difference in expected score if the athlete and the judge come from the same country.
+Hypothesis \#1 is tested by assessing the direction and magnitude of $B$.
 
-However, such a model does not take into account relevant differences between nations.
+Such a model does not take into account relevant differences between nations.
 A Norwegian judge may give higher scores to a fellow countryman’s performance due to a correspondence of their definitions of valuable stylistic elements, in accordance with national variations in evaluation practices.
@@
 This model will not confuse differences in stylistic culture with nationalistic biases, comparing the potentially biased judge against the four other judges for each jump.
-Following Zitzewitz (2006), we also estimated a model with so-called compensating biases, relaxing the assumption of objectivity with the four remaining judges.
-This model allows for the remaining judges to compensate a potentially biased judge, by lowering their scores when they know that one of the others will up his score.
-We did not find any significant compensating bias in our data, but this particular model did not converge and we thus abstain from reporting it in our set of results.
+The $B$ estimate recovered from this model will be compared with the corresponding parameter from model 1, and the result of this comparison can be applied to test hypothesis \#2.
 
@@
 If a judge consistently scores performances higher or lower than other judges, then we cannot be sure that the observed bias is not just a particular judge always scoring lower or higher than the average judge.
-To control for this, we add a judge fixed effect, $l_j$, which leads to our third model:
+To control for this, we add a judge fixed effect, $l_j$, which leads to an adjusted model \#2b:
 
@@
 This model is an improvement over the previous ones, but limitations still remain.
-The model assumes that bias is constant for each judge across all judges from all nations.
-Ideally, we would want to allow for variations in bias between individual judges or groups of judges by interacting the judge fixed-effect with the indicator function for judge and competitor nation correspondence.
-This would in principle let us obtain individual bias estimates for each judge, but our data unfortunately do not allow for such a specific estimation.
-They do however allow for estimating an interaction between nationalistic bias and the home country of the judge.
-This leads to the model:
+For example, there could be so-called compensating biases, where other judges on thethe same panel as a suspected biased judge compensate by lowering their scores when they know that one of the others will up his score.
+Finally, the model also assumes that bias is constant across all judges from all nations, a crucial assumption that we subsequently question in hypothesis \#3.
+
+## Variations across nations and time periods
+
+In order to test our third hypothesis, we want to allow for variations in bias between groups of judges from the same nation by interacting the judge fixed-effect in model \#2 with the indicator function for judge and competitor nation correspondence.
+This implies the followung model:
 
 \begin{equation}
-s_{ijp} = B∙(L_{j}∙\phi(I=J)) + q_p + \epsilon_{ijp}
+s_{ijp} = B_{j}∙(L_{j}∙\phi(I=J)) + q_p + \epsilon_{ijp}
 \end{equation}
 
-where $L_j$ is an added judge country fixed-effect. In this fourth model, the estimates for B provide an assessment of the average bias in scores by judges from different countries.
+where $L_j$ is an added judge country fixed-effect.
+The nation-specific estimates for B provide an assessment of the average bias in scores by judges from different countries.\footnote{In principle, we could obtain individual bias estimates for each judge and then analyze these as samples of judge biases scores grouped by judges' nationalities.
+This would be very demanding of our data and experiments indicate that estimates would be very imprecise.
+We thus refrain from doing this, and only allow the bias to vary between the different judge nationalities.} We also rank countries by our selected instrument of ski jumping history, namely the historical record of Olympic ski jumping gold medals in the decades before the seasons we study.
+The hypothesis is tested by estimating the Spearman rank correlation $\rho$ across the countries.
+A negative correlation indicates support for the hypothesis.
 
-## Estimation of home field advantage
-
-To analyse home field advantage we need to separate the problem of bias into the physical and the social domains.
-That is, whether an athlete actually performs better on the home field or whether he gets better scores by the judges due to social pressure or other factors.
-Thanks to the natural separation of style scores and length scores in ski jumping, this can be uncovered using similar models.
-
-In order to observe a home field effect in points, be it length points or the judges’ style points, we have to compare scores on a home field with scores on away fields – leading to the assumption that a jumper is somewhat consistent in performance throughout a season.
-Previously, we used the four other judges as a contrast measure of performance quality.
-For home field advantage we have to assume that an athlete maintains a relatively constant quality of jumping throughout the season, using his away jumps as a guide to his general performance level.
-This can be a bold assumption, given that local weather conditions and individual athlete’s “daily form” also may affect these scores, as well as the tiny margins of timing required to “push off” at the right moment.
-Another problem is that countries with a strong tradition for ski jumping will have both stronger jumpers, host competitions more often due to existing infrastructure, and have more judges in the judge panel.
-
-Again, we construct a dummy-variable using a function, $ω(I=H)$, which will take the value 1 if the jump takes place in country I, the athlete’s home country, and H being the competition country.
-We then wish to estimate the following model:
+Our fourth hypothesis regards potential changes in the variation of nationalist bias over time.
+In order to assess changes in the variation in nationalist bias, we estimate a model akin to model \#2, but with an added interaction with a dummy representing the second season (with the first season as a reference group).
+The parameter estimate for this interaction effect, $\Gamma$ provides a direct test of hypothesis with regard to changes in nationalist bias.
+The model is written as
 
 \begin{equation}
-s_{irs}' = \~B' ∙ω(I=H) + a_{i} +λ_{r} +\epsilon_{ir}
+s_{ijp} = B∙\phi(I=J) + \Gamma∙Y∙\phi(I=J)+ q_p + \epsilon_{ijp}
 \end{equation}
-
-where $s_{irs}$ are the length points of the athlete, in a specific season, in a specific competition.
-$\~B'$ is the estimated home field effect in points and $a_i$ is is a fixed effect for the athlete-season combinations.
-The fixed effect also removes any effect of individual jumpers that “strong jumpers” from traditionally strong ski jumping nations by out estimating each athlete, in each season, separately.
-Lastly, taking into account the different sizes of the hills, which may result in a skewness that the K-point system fails to adjust for, we add a fixed effect for the hill represented by $λ_r$.
-
-For the social part of the home field advantage we wish to estimate a similar model where style points awarded by the judges serves as the dependent variable.
-
-\begin{equation}
-s_{ijprs} = \~B ∙ ω(I=H) + a_{is} + λ_r + l_j + βx_p + \epsilon_{ijprs}
-\end{equation}
-
-In this model, we add fixed effects for judges as in (3) to capture leniency of individual judges.
-We cannot use jump fixed-effects to capture the inherent objective quality of the jump, because home advantage is invariant within a jump, but we can include two central jump characteristics in our data, namely length and speed.
-This is because “all else being equal, a ski jumper with the same take-off speed but better form while airborne will travel further” [@zitzewitz_nationalism_2006, 78].
-We therefore add these two variables to the model, where they and their coefficients are represented by the $βx_p$.
-
-In accordance with norms of open social science, we offer our data and analysis log in an open repository at the time of publication.
 
@@
 
+Our analysis of nationalistic bias proceeds in four steps, with each step corresponding to a test of one of our four hypotheses.
+
 ## Nationalistic biases in judge scores
 
-Our analysis of nationalistic bias proceeds in four steps. Table 2 presents the results from the first three estimations.
+Table 2 presents the results from the first three estimations.
+Our first model yields a positive and statistically significant estimate: judges award an average of 0.146 extra points to jumpers from their home country.
+This is approximately one-third of the minimum increment in judge scores in ski jumping (0.5).
+Compared with within-jump variance it only amounts to a quarter of a standard deviation.
+This result already suggests that the magnitude of nationalistic bias in judges’ scores is limited.
+The estimate may be biased by, for example, cultural and historical differences between countries that lead to an over- or underappreciation of certain styles.
 
@@
 ```{r tab2, echo=FALSE, results='asis'}
-stargazer(model1, model2, model3, out.header=FALSE, header=FALSE,
+stargazer(model1, model2, model2b, out.header=FALSE, header=FALSE,
     single.row=TRUE,
-    column.labels = c("Model 1a: Naïve", "Model 2a: Within-jump", "Model 3a: Within-judge"),
+    column.labels = c("Model 1: Naïve", "Model 2a: Within-jump", "Model 3a: Within-judge"),
     title="Results from regression models of national bias",
@@
     notes.align="l",
-    notes = c("In models 2a and 3a, the intercept is suppressed by the fixed effects estimation procedure."),
+    notes = c("In models 2 and 2b, the intercept is suppressed by the fixed effects estimation procedure."),
     notes.append=TRUE,
@@
 
-\begin{landscape}
-```{r tab2b, echo=FALSE, results='asis'}
-stargazer(model1b, model2b, model3b, out.header=FALSE, header=FALSE,
-    single.row=TRUE,
-    column.labels = c("Model 1b: Naïve", "Model 2b: Within-jump", "Model 3b: Within-judge"),
-    title="Results from regression models of national bias for season 2015-2016",
-    covariate.labels = c("National bias"),
-    dep.var.caption  = "", dep.var.labels.include = FALSE,
-    model.numbers = FALSE, model.names = FALSE,
-    add.lines = c(" "),
-    notes.align="l",
-    notes = c("In models 2b and 3b, the intercept is suppressed by the fixed effects estimation procedure."),
-    notes.append=TRUE,
-    df = TRUE, decimal.mark =".",
-    ci=TRUE, ci.level=0.95, ci.separator = "  ",
-    omit.stat=NULL,
-    star.cutoffs = c(0.05,0.01,0.001)
-    )
-```
-\end{landscape}
-
-Our first model yields a positive and statistically significant estimate: judges award an average of 0.146 extra points to jumpers from their home country.
-This is approximately one-third of the minimum increment in judge scores in ski jumping (0.5).
-Compared with within-jump variance it only amounts to a quarter of a standard deviation.
-This result already suggests that the magnitude of nationalistic bias in judges’ scores is limited.
-The estimate may be biased by cultural and historical differences between countries that lead to an over- or underappreciation of certain styles.
-	The second model takes up exactly this problem, and provides a within-jump estimate of the nationalistic bias.
+This was the concern of the second hypothesis, were we assumed that the bias level would decrease once we controlled for other, unobserved characteristics of the jump.
+The second model takes up exactly this challenge, and provides a within-jump estimate of the nationalistic bias.
 In this model, the coefficient is estimated at 0.068, markedly lower than in the first model.
@@
 The magnitude is low, but there is clearly a tendency towards nationalistic bias.
+In our adjusted second model \#2b, where we also include measures of individual judges' leniency, we obtain a comparable estimate of nationalistic bias.
 
-The third model let us also include judge leniency, a measure of how strict a judge is in his or her evaluations.
-From this model, we obtain a similar estimate of nationalistic bias of 0.081.
-All estimates are well below 0.5, the minimum increment in judge scores used in ski jumping.
-Judges’ leniency display important variation, however, as the range of judges leniency coefficients is larger than 0.8 points.
-Figure 1 show this variation in a histogram.
-\bigskip
+## Variations in bias across nations and time periods
 
-```{r fig1, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=5, fig.align="center"}
-judge_effects <- filter(as.data.frame(getfe(model3, se = FALSE)),fe=="judge_name")
-figure1 <- ggplot(data=judge_effects, aes(x=effect)) + 
-    geom_histogram(bins=15) +
-    xlab("Leniency") + ylab("Number of judges") +
-    theme_minimal() +
-    ggtitle("Figure 1. Histogram of judges' leniency")  
-#figure1
+Our third hypothesis concerned variation in nationalistic biases for judges from different countries, and whether such variation is correlated with an indicator of the historical importance of ski jumping in that country.
 
-judge_effectsb <- filter(as.data.frame(getfe(model3b, se = FALSE)),fe=="judge_name")
-figure1b <- ggplot(data=judge_effectsb, aes(x=effect)) + 
-    geom_histogram(bins=15) +
-    xlab("Leniency") + ylab("Number of judges") +
-    theme_minimal() +
-    ggtitle("Figure 1. Histogram of judges' leniency")  
-#figure1b
-
-botheffects <- list("2005-2007" = judge_effects, "2015-2016" = judge_effectsb) %>% 
-    map(function(df) select(df, effect)) %>% 
-    bind_rows(.id="Season")
-figure1c <- ggplot(data=botheffects, aes(x=effect, fill=Season)) + 
-    geom_histogram(bins=20, alpha=0.5) +
-    xlab("Leniency") + ylab("Number of judges") +
-    theme_minimal() +
-    theme(legend.position = "bottom") + 
-    ggtitle("Figure 1. Histogram of judges' leniency")  
-figure1c
-```
-\bigskip
-
-The fourth and final model allowed estimating the variation in nationalistic biases for judges from different countries.
-The results are shown in Figure 2.
-Countries with fewer than 25 observations are not shown, as their estimates have large standard errors and thus are imprecisely estimated.
-French judges appear most biased, followed by Polish, and Japanese judges.
-For the other countries, the estimated biases are fairly small and in several cases not statistically significant.
-It is difficult to provide a substantive interpretation of this variation in national biases, but it is nevertheless clear that the biases are relatively small in most cases.
+Our third model allowed estimating such variation, and the results are shown in Figure 2.
+Judges from Russia, the United States and France appear most biased.
+Norwegian and Slovakian judges seems to hold a negative nationalistic bias, where same-nation jumpers are punished more severely than others.
+Overall, the estimated biases seem fairly small and in several cases not statistically significant.
+It is difficult to provide a substantive interpretation of this variation in national biases, but it is nevertheless clear that the biases are relatively small in magnitude.
+The caseof Slovakia is curious, as this country show particularly strong negative bias.
+The estimate lie well outside of the rest of the estimates, and we will treat it as an outlier in the following analysis.
 
@@
 ```{r fig2, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=5, fig.align="center"}
-nationbias <- broom::tidy(model4) %>% 
+nationbias <- broom::tidy(model3) %>% 
     mutate(country = str_sub(term, start=-3, end=-1)) %>% 
-#    filter(!(country %in% c("USA","SVK","SWE","ITA"))) %>%
     mutate(ocountry = factor(country, levels = .$country[rev(order(.$estimate))])) %>% 
@@
     ggtitle("Figure 2. Nation-specific estimates of judges' bias")
-#figure2
+figure2
 
-nationbiasb <- broom::tidy(model4b) %>% 
-    mutate(country = str_sub(term, start=-3, end=-1)) %>% 
-    mutate(ocountry = factor(country, levels = .$country[rev(order(.$estimate))])) %>% 
-    mutate(lowci = estimate-1.96*std.error, hici=estimate+1.96*std.error) %>% 
-    arrange(estimate)
-
-figure2b <- ggplot(data = nationbiasb,mapping = aes(x = ocountry, y = estimate)) +
-    geom_point(size=3) +  
-    geom_pointrange(aes(x=country, y=estimate, ymin = lowci, ymax=hici), fatten = 1.2) +
-    #color="grey") +
-    #geom_segment(aes(x=country, y=estimate, xend=country, yend=0)) +
-    xlab("Nation") + ylab("Strength of bias") +
-    theme_minimal() +
-    ggtitle("Figure 2. Nation-specific estimates of judges' bias")
-#figure2b
-
-bothbias <- list("2005-2007" = nationbias, "2015-2016" = nationbiasb) %>% 
-    map(function(df) select(df, country, estimate, lowci, hici)) %>% 
-    bind_rows(.id="Season") %>% 
-    arrange(Season, estimate)
-figure2c <- ggplot(data = bothbias,mapping = aes(x = forcats::fct_rev(forcats::fct_reorder(country, estimate)), y = estimate, color=Season)) +
-    geom_point(size=3) +  
-    geom_pointrange(aes(x=country, y=estimate, ymin = lowci, ymax=hici), fatten = 1.2) +
-    xlab("Nation") + ylab("Strength of bias") +
-    theme_minimal() +
-    theme(legend.position = "bottom") + 
-    ggtitle("Figure 2. Nation-specific estimates of judges' bias")
-figure2c
 ```
@@
 
-### Home field advantages
+To properly test the third hypothesis, we then estimate the rank correlation of the nation-specific bias coefficient and the nation's Olympic medal harvest.
+This exercise was done by scraping existing public statistics on Olympic medals, and calculating how many medals each country has collected.
 
-Table 3 presents the results for the two models, where we included effects of the hypothesized home field advantage.
-Following the distinction between the physical and social dimensions of such advantages, we specified one model for the actual length of the jump and another for style scores.
-The first column of Table 3 show the parameter estimates for the length model (Model 5).
-This model includes competitor-season-fixed effects, which implies that variations among different athletes and their form is taken into account and the estimate for differential jump lengths for home field competitors are purged of those influences.
-Our main interest lies with the dummy variable for the athlete performing in his home country.
-The parameter estimate for this dummy variable is not significantly different from zero.
-Athletes thus do not seem to perform better in their home countries, than when competing in other countries.
-The physical dimension of home field advantage is thus undetectable in our data.
+```{r scrape-gold, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
+library(rvest)
+theurl <- "https://en.wikipedia.org/wiki/List_of_Olympic_medalists_in_ski_jumping"
+webpage <- read_html(theurl)
 
-\begin{landscape}
-```{r tab3, echo=FALSE, results='asis'}
-stargazer(model5, model6, out.header=FALSE, header=FALSE,
-    single.row=TRUE,  omit = c("judge","competitionplace"), 
-    column.labels = c("Model 5: Length points", "Model 6: Style points"),
-    title="Results from regression models of home field advantage", 
-    covariate.labels = c("Home field"),
-    dep.var.caption  = "", dep.var.labels.include = FALSE,
-    model.numbers = FALSE, model.names = FALSE,
-    add.lines = c(" "),
-    notes.align="l",
-    notes = c("Both models include athlete and jumping hill fixed effects, and model 6 also include judge fixed-effects."),
-    notes.append=TRUE,
-    df = TRUE, decimal.mark =".",
-    ci=TRUE, ci.level=0.95, ci.separator = "  ",
-    omit.stat=NULL,
-    star.cutoffs = c(0.05,0.01,0.001)
-    )
+lh_table <- webpage %>% 
+    html_nodes("table:nth-child(6)") %>% 
+    html_table(header=T)
+
+nh_table <- webpage %>% 
+    html_nodes("table:nth-child(11)") %>% 
+    html_table(header=T)
+
+goldindex <- bind_rows(lh_table, nh_table) %>% 
+    mutate(year = as.numeric(str_sub(Games,start=1,end=4))) %>%
+    filter(year<2006) %>% 
+    select(-Games) %>% 
+    gather(key=type, value=winner, -year) %>% 
+    filter(winner!="not awarded") %>% 
+    mutate(country = stringi::stri_extract_last_words(winner)) %>% 
+    mutate(country = case_when(country=="Union" ~ "RUS",
+                               country=="States" ~ "USA",
+                               country=="Japan" ~ "JPN",
+                               country=="Germany" ~ "GER",
+                               country=="Austria" ~ "AUT",
+                               country=="Switzerland" ~ "SUI",
+                               country=="Finland" ~ "FIN",
+                               country=="Sweden" ~ "SWE",
+                               country=="Poland" ~ "POL",
+                               country=="Norway" ~ "NOR",
+                               country=="Yugoslavia" ~ "SLO",
+                               country=="Czechoslovakia" ~ "SVK",
+                               TRUE~country)) %>% 
+    group_by(country) %>% 
+    summarize(medals = n()) %>% 
+    arrange(medals)
+cze <- goldindex %>% filter(country=="SVK") %>% pull(medals)
+goldindex <- goldindex %>% add_row(country="CZE", medals=cze)
+
+ranking <- left_join(goldindex, nationbias, by="country")
+
+figure3 <- ggplot(ranking, aes(x=medals, y=estimate, label=country)) + 
+    geom_point(size=2) + 
+    geom_smooth(method="lm", se=F) +
+    geom_text(nudge_x = 1) +
+    theme_minimal() +
+    labs(x="Olympic medals up to 2002", y="Nationalist bias estimate", title="3a. Olympic medals and bias estimates")
+
+figure3b <-ggplot(filter(ranking, country!="SVK"), aes(x=medals, y=estimate, label=country)) + 
+    geom_point(size=2) + 
+    geom_smooth(method="lm", se=F) +
+    geom_text(nudge_x = 1) +
+    theme_minimal() +
+    labs(x="Olympic medals up to 2002", y="Nationalist bias estimate", title="3b. Outlier removed")
+
+spearman <- cor.test( ~ medals + estimate, 
+         data=ranking,
+         method = "spearman",
+         continuity = FALSE,
+         conf.level = 0.95)
+
 ```
-\end{landscape}
 
-We examine the social dimension of home field advantage in Model 6, where we include a parameter for bias in judge's scores if the athlete is competing in his or her home country.
-If there is home field advantage, this parameter should be measurably larger than zero.
-The estimate of this parameter is statistically significant, and larger than zero, suggesting that there is a social dimension of home field advantage in ski jumping.
-The effect is small and roughly of the same magnitude (0.074) as for nationalistic bias.
-The bias only amounts to a quarter of a standard deviation within each jump and it is well below the minimum increment for judge scores of 0.5 point.
-It seems unlikely that home field advantages is a serious threat to the validity of results of any competition.
+The Spearman Rho rank correlation coefficient of nationalist bias and Olympic medals was estimated at ```spearman$estimate``` which thus supports the hypothesis that the magnitude of bias is negatively related to the historical importance of ski jumping in the country (as measured by our index of Olympic medals).
+
+Figure 3 shows the scatterplot between Olympic medals and natonalistic bias estimates, with and without the outlier Slovakia included.
+\smallskip
+
+```{r figthree, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+library(patchwork)
+
+figure3 + figure3b + plot_layout(ncol = 2, widths = c(1,1))
+
+```
 
@@
 
-We conducted separate analyses for home-field bias in performance (as indicated by the length of the jumps) and in judges’ evaluations of the stylistic elements in each jump.
-The results showed no effects on the jumpers’ actual performance, implying that the jumpers did not benefit from familiarity with the hill or from home-audience support in terms of jump length.
-Instead, home-field bias showed up in the judges evaluations of the jumps, even though the effect was here again of very small magnitude.
-Whether the bias stemmed from “crowd-pleasing”—which functions as a conflicting motivation to objectivity in the judges’ performance—or some other bias-generating mechanism is beyond the scope of this study.
-
 As has been pointed out in previous research, biases in evaluations need not result from conscious actions by the judges.
@@
 
-We estimated nationalistic and home-field advantages in ski jumping using data from the entire FIS World Cup 2006/2007 and 2007/2008 seasons.
-From our results, we confirm parts of the extant literature, and document both national and home field biases in judges’ evaluations.
+In this paper, we report an analysis of data from the entire FIS World Cup 2006/2007 and 2007/2008 seasons in Nordic ski jumping.
+Our overarching research question is whether there is nationalistic biases in ski jumping, and whether there are systematic variations in such bias when influences of the actual performance are ruled out.
+From our results, we confirm a tenet of the extant literature, and document nationalistic biases in judges’ evaluations.
 The biases were generally small, and do not seem to pose a major threat to the validity of results from skijumping competitions.
 However, in league or series-type competition tournaments, the cumulative effects of nationalistic bias may be important.
-An important takeaway from this study is that the system of regulations in
+An important takeaway from this study is that the system of regulations in ski jumping that is already in place most likely provide adequate mitigation of judges' nationalist impulses so that the objectivity and fairness of the results are not threatened.
 

### 2019-08-12 revised-manuscript/home-field-snippets.R

@@ -1,0 +1,90 @@
+
+# HOME FIELD BIAS MODELS WITH 2016 DATA
+# 
+home16 <- season2016 %>% 
+    select(competitorName, place, country, competitorNation, lenpts = lengthPoints) %>% 
+    mutate(competitionplace = str_sub(place, start=1,end=-6)) %>% 
+    mutate(athome = if_else(country==competitorNation,1,0))
+
+# Model 5: Length model
+model5b <- felm(data=home16, lenpts ~ athome + competitionplace | competitorName)
+summary(model5b)
+
+# Model 6: Style points model
+model6b <- felm(data=scores16, judge ~ athome + competitionplace + judge_name + speed + length| competitorName)
+summary(model6b)
+
+
+## The extant literature on home field advantages
+
+Home field advantage is another type of potential bias in ski jumping. Such advantages, where some competitors benefit from advantages from competing on the home field, may arise in many different ways. Several of the mechanisms that may produce home field advantages are pscyhological sources of superior performances, while others involve judges overrating home field competitors. This point illustrates that the mere presence of a biases is not necessarily illegitimate. In order to improve scoring and fairness in sports, one must understand the mechanisms behind them. 
+
+Some research into the home-field advantage has attempted to move our understanding of the causal mechanisms behind  the advantages. Carmichael and Thomas [-@carmichael_homefield_2005] outline the four most common explanations of this phenomena using soccer as an example. The first is a familiarity explanation – the home team is more familiar with the field, the stadium, the locker room, and so forth, which puts the players more at ease in the situation, allowing them to concentrate on performing at their best. Second, traveling to another field could result in fatigue, invoke various stressors, and cause disruptions in pre-competition routines for the away team. All these factors may lead to inferior performances by the away team, relative to their likely performance at the home field. Third, specific rules may favour the home team, giving them the upper hand. Last, there could be a crowd effect. The size of the crowd at the home field might improve individual or team performance. It might also make the judge to be more lenient for the home team, through psychological mechanisms.
+
[…]

### 2019-08-14 revised-manuscript/response_to_reviewers.Rmd
Revised manuscript ready

@@
 
-We now build upon sociological theory on evaluation to propose hypotheses on variation in biases in the field of ski jumping.
+We now build more explicitly upon sociological theory on evaluation to propose hypotheses on variation in biases in the field of ski jumping.
 
@@
 
-We are grateful to the Editor for her/his comments on our manuscript. They helped us writerethink and rewrite the work in a way that we believe is a stronger contribution to the field. Thank you!
+We are grateful to the Editor for her/his comments on our manuscript. They helped us rethink and rewrite the work in a way that we believe is a stronger contribution to the field. Thank you!
 
@@
 
-Our estimate from model 1, the "naïve" model, may be overestimating the strength of bias for such reasons, and in model 2 we control for *all* unobserved characteristics of the jump that affects the judge scores.
+Our estimate from model 1, the "naïve" model, may be overestimating the strength of bias for such reasons, and in model 2 we control for *all* unobserved characteristics of the jump that affects the judge scores by way of jump fixed-effects.
 
@@
 Yes, we use a fixed-effects term for each jump to remove confounding factors from characteristics of the jump.
-However, we believe this model does _not_ rely on an assumption of unbiasedness of the other four scores nor an assumption that there is only one judge from the competitor's nation in the judge panel.
+
+However, we believe this model does _not_ rely on an that there is only one judge from the competitor's nation in the judge panel.
 If there are multiple judges from the competitor's country, there will simply be multiple 1's among the five observations that make up the observations of judge scores for each jump.
@@
 As we have quite a few of such performances, we are able to identify and estimate an effect of nationalistic bias on judge scores.
+
+The assumption of unbiasedness among other panel judges can clearly be questioned, and it is possible to estimate counteracting "compensating biases" (á la Zitzewitz 2006).
+Our experiments did not find that results depended on inclusion of such compensating biases.
 

### 2019-08-14 revised-manuscript/manuscript.Rmd
Revised manuscript ready

@@
 keywords: nationalist bias, skijumping, judges, fixed effects
-abstract: "High stakes and prizes in international sports competitions necessitate efficiently identifying legitimate winners.
-Illegitimate biases of non-independent judges represent a major threat to the legitimacy of international competitions.
-Judges’ nationalities represent one source of bias, as they may prefer athletes of their own nation.
-Judges may also favour athletes competing within their home countries, creating another potential bias.
-Such biases, conscious or unconscious, may affect the outcomes of competitions.
-The literature offers no complete consensus on the magnitude of such biases, hence not whether they affect competition outcomes or not.
+abstract: "Illegitimate biases of non-independent judges represent a threat to the legitimacy of international sport competitions.
+Judges’ nationalities represent one source of bias, as they consciously or subconsciously may prefer athletes from their own nation.
+Such biases may affect the outcomes of competitions.
+The literature offers no complete consensus on the magnitude, origin and stability of such biases.
 In this paper, we shed light on these problems, using international ski jumping as an example.
-We draw on data from the FIS World Cup competitions in the 2006-2008 and 2015-2016 seasons, and employ several fixed-effects based regression designs.
-Our results reaffirm suspicions of national bias in major ski jumping competitions, but also show their magnitude is too little to be of major relevance to competition outcomes.
-There is a trend towards smaller nationalist biases over time, which suggests that increasing professionalization of sports also may limit the influence of subjective and nationalist impulses."
+We draw on data from the FIS World Cup competitions in the 2006-2008 and 2015-2016 seasons and estimate a series of fixed-effects models to test hypotheses on nationalist biases.
+Our results reaffirm suspicions of nationalist bias in major ski jumping competitions, but also show their magnitude is too little to be of major relevance to competition outcomes.
+The biases vary between nations, but do not change markedly over the study period.
+"
 thanks: "The authors are grateful for financial support from the Småforsk programme at the University of Oslo, and to Marte Knutsson for research assistance in the early phase of establishing the data set.
@@
 library(ggthemes)
+library(ggrepel)
 library(here)
@@
 
-# Model 4a: Estimating changes over time
-model4a <- felm(data = all_scores, judge ~ natbias + natbias*seasondummy + seasondummy| jumpid)
-summary(model4a)
-
-# Model 4b: Estimating changes in leniency and national variance
-
-## Alternative model 4 with individual-level bias estimates. Not included in the final paper.
-model4_indjudges <- felm(data=all_scores, formula=judge ~  natbias*judge_name | jumpid) 
-model4_indjudges %>% 
-    tidy %>% 
-    mutate(natbias = if_else(str_detect(term,"natbias"),"interaction","main")) %>% 
-    mutate(judge_name = str_sub(term, start=1+((natbias=="interaction")*8))) %>% 
-    mutate(judge_name = if_else(judge_name=="", "overall",judge_name)) %>% 
-    select(natbias, judge_name, estimate) %>% 
-    group_by(judge_name) %>% 
-    spread(key=natbias, estimate) %>% 
-    rename(bias = interaction, leniency = main) %>% 
-    ungroup %>% 
-    mutate(judge_name = str_sub(judge_name,start=11)) %>% 
-    head
+# Model 4: Estimating changes over time in national bias
+model4 <- felm(data = all_scores, judge ~ natbias + natbias*seasondummy + seasondummy| jumpid)
+summary(model4)
+bias_main <- tidy(model4) %>% filter(term=="natbias") %>% pull(estimate)
+bias_interact <- tidy(model4) %>% filter(term=="natbias:seasondummy") %>% pull(estimate)
 
@@
 They prefer their own to win, and give higher scores to those athletes.
+Thus crude form of nationalist bias is similar to what has denoted taste-based discrimination, which refers to individuals’ preferences for certain categories of people and is often used to explain discrimination against women or ethnic minorities in the labour market [@becker1957economics].
+
+Applied to ski jumping, the mechanism refers to judges’ conscious nationalistic impulses that bias their evaluations in favour of their own, and against competitors.
+A subtler form of this mechanism leads to subconscious bias for jumpers from one’s own country, where own preferences or social expectations lead judges to biased evaluations [@bassett2005rating, @campbell1996nonparametric].
 Importantly, this theoretical mechanism does not by itself predict variation in the magnitude of biases across countries.
-It could in princple be combined with other theoretical ideas to arrive at predictions about any such variation.
+It can be combined with other theoretical ideas to arrive at predictions about any such variation.
 
@@
 Eventually he won competitions, due to the technical advantages of the V-style, and other actors started to mimic the style [@muller_performance_2008].
-These changes then led to a change in the room for evaluations, where V-style was accepted and it is currently completely dominant in the field.
+These changes then led to a change in the room for evaluations, where V-style was accepted and it is currently completely dominant in the field [@pfister2007sportification].
 
@@
 For the season 2015-2016, we wrote automated web scraping software that identified relevant PDF files on the FIS servers, downloaded these files, and scraped them for results according to our pre-specified instructions.
-In accordance with norms of open social science, the software used to download and scrape the PDF files are documented in the GitHub repository associated with this article.[^1: See https://github.com/torkildl/skijumping for data and software.]
+In accordance with norms of open social science, the software used to download and scrape the PDF files are documented in the GitHub repository associated with this article.
 
@@
 
-Such a model does not take into account relevant differences between nations.
+Such a model does not take into account other relevant differences between nations.
 A Norwegian judge may give higher scores to a fellow countryman’s performance due to a correspondence of their definitions of valuable stylistic elements, in accordance with national variations in evaluation practices.
 
-Our next step is an attempt to remove such confounding factors, by exploiting the other judges’ evaluations of the same jump in estimation of nationalistic bias.
+Our next step is an attempt to remove such factors, by exploiting the other judges’ evaluations of the same jump in estimation of nationalistic bias.
 Assuming that the other four judges give an unbiased score, a more precise way to estimate the bias is to add a fixed effects-term for each jump to the model (in effect including a dummy variable for each jump).
@@
 This model is an improvement over the previous ones, but limitations still remain.
-For example, there could be so-called compensating biases, where other judges on thethe same panel as a suspected biased judge compensate by lowering their scores when they know that one of the others will up his score.
+For example, there could be so-called compensating biases, where other judges on the same panel as a suspected biased judge compensate by lowering their scores when they know that one of the others will up his score.
 Finally, the model also assumes that bias is constant across all judges from all nations, a crucial assumption that we subsequently question in hypothesis \#3.
@@
 This would be very demanding of our data and experiments indicate that estimates would be very imprecise.
-We thus refrain from doing this, and only allow the bias to vary between the different judge nationalities.} We also rank countries by our selected instrument of ski jumping history, namely the historical record of Olympic ski jumping gold medals in the decades before the seasons we study.
+We thus refrain from doing this, and only allow the bias to vary between the different judge nationalities.} We also rank countries by our selected instrument of ski jumping history, namely the historical record of Olympic ski jumping gold medals in the decades up to and including 2002 (before the seasons we study.
+This medal variable was constructed by scraping existing public statistics on Olympic medals in ski jumping competitions (men's Large Hill and Normal Hill only), and calculating how many medals each country has collected.
 The hypothesis is tested by estimating the Spearman rank correlation $\rho$ across the countries.
@@
 In order to assess changes in the variation in nationalist bias, we estimate a model akin to model \#2, but with an added interaction with a dummy representing the second season (with the first season as a reference group).
-The parameter estimate for this interaction effect, $\Gamma$ provides a direct test of hypothesis with regard to changes in nationalist bias.
+The parameter estimate for this interaction effect, $\gamma$ provides a direct test of hypothesis with regard to changes in nationalist bias.
 The model is written as
@@
 \begin{equation}
-s_{ijp} = B∙\phi(I=J) + \Gamma∙Y∙\phi(I=J)+ q_p + \epsilon_{ijp}
+s_{ijp} = B_{t_1}∙\phi(I=J) + \gamma∙B_{t_2}∙\phi(I=J)+ q_p + \epsilon_{ijp}
 \end{equation}
+
+In principle, this model could include also a general tendency in judge leniency over time, but this term will be subsumed due to perfect linearity with sets of jump fixed effects (as specific jumps only take place in one season).
 
@@
 
-Our analysis of nationalistic bias proceeds in four steps, with each step corresponding to a test of one of our four hypotheses.
+Our analysis of nationalistic bias proceeds in several steps.
+We first assess the first two hypotheses on the mere existence of nationalistic biases, and subsequently move on to the last two hypotheses on variation across nations and time periods.
 
@@
 Table 2 presents the results from the first three estimations.
-Our first model yields a positive and statistically significant estimate: judges award an average of 0.146 extra points to jumpers from their home country.
-This is approximately one-third of the minimum increment in judge scores in ski jumping (0.5).
+Our first model yields a positive and statistically significant estimate: judges award an average of 0.136 extra points to jumpers from their home country.
+This is nearly one-third of the minimum increment in judge scores in ski jumping (0.5 points).
 Compared with within-jump variance it only amounts to a quarter of a standard deviation.
@@
     single.row=TRUE,
-    column.labels = c("Model 1: Naïve", "Model 2a: Within-jump", "Model 3a: Within-judge"),
+    column.labels = c("Model 1: Naïve", "Model 2a: Within-jump", "Model 2b: Within-judge"),
     title="Results from regression models of national bias",
@@
     notes.align="l",
-    notes = c("In models 2 and 2b, the intercept is suppressed by the fixed effects estimation procedure."),
+    notes = c("In models 2a and 2b, the intercept is suppressed by the fixed effects estimation procedure."),
     notes.append=TRUE,
@@
 This was the concern of the second hypothesis, were we assumed that the bias level would decrease once we controlled for other, unobserved characteristics of the jump.
-The second model takes up exactly this challenge, and provides a within-jump estimate of the nationalistic bias.
-In this model, the coefficient is estimated at 0.068, markedly lower than in the first model.
+Model 2b is a response to exactly this challenge, and provides a within-jump estimate of the nationalistic bias.
+In this model, the coefficient is estimated at 0.065, markedly lower than in the first model.
 It is statistically significant at the 0.1% level.
 The magnitude is low, but there is clearly a tendency towards nationalistic bias.
-In our adjusted second model \#2b, where we also include measures of individual judges' leniency, we obtain a comparable estimate of nationalistic bias.
+In our adjusted second model \#2b, where we also include measures of individual judges' leniency, we obtain a comparable estimate of nationalistic bias (0.075, sig.
+at 0.1\% level).
+
+The conclusion thus far is that there are nationalistic biases in evaluations made by judges in the FIS World Cup competitions over the last decades.
+These biases do not disappear once we control rigorously for unobserved characteristics of both the jump itself and the judges.
+The first two hypotheses are supported by the data analysis.
+We do not know from these results, whether the remaining nationalistic bias is due to the social psychological, cultural-historical or the differential-professionalization mechanisms.
+It seems fair to allow all three to play a role.
 
@@
 It is difficult to provide a substantive interpretation of this variation in national biases, but it is nevertheless clear that the biases are relatively small in magnitude.
-The caseof Slovakia is curious, as this country show particularly strong negative bias.
-The estimate lie well outside of the rest of the estimates, and we will treat it as an outlier in the following analysis.
+
+The case of Slovakia is curious.
+This country's judges show a particularly strong negative bias.
+As the estimate lie well outside of the rest of the sample, it may well be reasonable to treat it as an outlier in the following analysis.
 
@@
 
-To properly test the third hypothesis, we then estimate the rank correlation of the nation-specific bias coefficient and the nation's Olympic medal harvest.
-This exercise was done by scraping existing public statistics on Olympic medals, and calculating how many medals each country has collected.
+By itself, this piece of evidence is not enough to test hypothesis \#3, as it also suggests several specific mechanisms for the variation in biases, namely cultural-historical legacies and differential professionalization.
+According to these mechanisms, the bias should be lower in countries where ski jumping has a stronger historical role in sports and where the degree of professionalization is the highest.
+We are unable to directly adjudicate between these two mechanisms, as there is likely a strong positive correlation between the place of ski jumping in the national sports field and the degree of professionalization.
+
+To test the third hypothesis, we estimate the rank correlation of the nation-specific bias coefficient and the nation's Olympic medal harvest.
 
@@
 
-figure3 <- ggplot(ranking, aes(x=medals, y=estimate, label=country)) + 
+figure3 <- ggplot(filter(ranking, country!="SVK"), aes(x=medals, y=estimate, label=country)) + 
     geom_point(size=2) + 
-    geom_smooth(method="lm", se=F) +
-    geom_text(nudge_x = 1) +
-    theme_minimal() +
-    labs(x="Olympic medals up to 2002", y="Nationalist bias estimate", title="3a. Olympic medals and bias estimates")
-
-figure3b <-ggplot(filter(ranking, country!="SVK"), aes(x=medals, y=estimate, label=country)) + 
-    geom_point(size=2) + 
-    geom_smooth(method="lm", se=F) +
-    geom_text(nudge_x = 1) +
-    theme_minimal() +
-    labs(x="Olympic medals up to 2002", y="Nationalist bias estimate", title="3b. Outlier removed")
+    geom_smooth(method="lm", se=F, color="brown") +
+    geom_text_repel() +
+    xlab("Olympic medals") + ylab("Strength of bias") +
+    ggtitle("Figure 3. Strengh of bias and Olympic medals") + 
+    theme_minimal()
 
@@
 
-The Spearman Rho rank correlation coefficient of nationalist bias and Olympic medals was estimated at ```spearman$estimate``` which thus supports the hypothesis that the magnitude of bias is negatively related to the historical importance of ski jumping in the country (as measured by our index of Olympic medals).
+The Spearman $\rho$ rank correlation coefficient of nationalist bias and Olympic medals was estimated at ```r round(spearman$estimate,3)``` which thus supports the hypothesis that the magnitude of bias is negatively related to the Olympic medal harvest for the country.
+If we can assume that historical medal harvest is postively related to the relative place of ski jumping in the national sports field and/or the degree of professionalization of ski jumping, our third hypothesis must be considered supported by the data.
 
-Figure 3 shows the scatterplot between Olympic medals and natonalistic bias estimates, with and without the outlier Slovakia included.
+Figure 3 shows the scatterplot between Olympic medals and natonalistic bias estimates.
+An identical scatterplot that also includes the outlier Slovakia, would yield a slightly less negative relationship.
+
 \smallskip
 
-```{r figthree, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-library(patchwork)
-
-figure3 + figure3b + plot_layout(ncol = 2, widths = c(1,1))
-
+```{r figthree, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=5, fig.align="center"}
+figure3
 ```
 
+We now turn to our fourth hypothesis, which stated that nationalistic biases would decrease over time.
+We test this hypothesis by estimating a fourth model that includes an interaction coefficient with time period.
+We thus obtain two coefficients for national bias: a main effect and an interaction effect.
+The latter quantifies the change in bias from the 2006-2008 seasons to the 2015-2016 season.
+\newpage
+
+```{r tab3, echo=FALSE, results='asis'}
+stargazer(model4, out.header=FALSE, header=FALSE,
+    single.row=TRUE,
+    column.labels = c("Model 4: Change over time"),
+    title="Results from regression models of change over time in nationalistic bias",
+    covariate.labels = c("Main bias effect","","Interaction with time"),
+    dep.var.caption  = "", dep.var.labels.include = FALSE,
+    model.numbers = FALSE, model.names = FALSE,
+    add.lines = c(" "),
+    notes.align="l",
+    notes = c("The intercept is suppressed by the fixed effects estimation procedure."),
+    notes.append=TRUE,
+    df = TRUE, decimal.mark =".",
+    ci=TRUE, ci.level=0.95, ci.separator = "  ",
+    omit.stat=NULL,
+    star.cutoffs = c(0.05,0.01,0.001)
+    )
+```
+
+As can be seen from Table 3, the interaction effect is not measurably different from zero.
+It is estimated at  ```r round(bias_interact,3)``` which is a miniscule magnitude relative to the main effect of ```r round(bias_main,3)```.
+All in all, there does not seem to be much change in nationalistic biases over the two time periods our data set covers.
+The fourth hypothesis must thus be rejected.
+This implies that if there is an ongoing professionalization of the sport, it has not resulted in lower nationalistic biases among judges.
+
 # Our results in light of the extant literature
+
+In sum, we have learned that there are nationalistic biases in ski jumping evaluations. These biases persist when controlling for characteristics of the performance and individual judges.
+The bias is less strong among nations with a historical culture of ski jumping. The bias have not declined over a decade of potentially increasing professionalization.
+
+What can we learn from these results in totality?
+At the very least, we can speculate about the origin of nationalistic biases.
+The simplistic social pscyhological mechanism is supported by the data, by the mere existence of biases.
+However, as the social psychology should not be very different across these nations, the variation among nations remains a puzzle.
+The absence of changes over time in bias strength is weak evidence against the idea that biases are results of "unprofessional practices" and will disappear once professionalization increases.
+A possibility that remains is that, as biases are stronger among nations where ski jumping is less important in the national sports field, judges (unconsciously) from such nations give higher scores to up the chance of notable, positive results and support their sport on the national scene.
 
@@
 If this variation is considered to be random deviations from a “true” score, it is clear that competitors should worry more about chance variations in judges’ evaluations (and possible idiosyncratic opinions) than about nationalist biases affecting the outcome of the competition.
+The ski jumping sport has by implementing several safeguards reduced the importance of nationalistic biases to a minimum.
 
@@
 
-In this paper, we report an analysis of data from the entire FIS World Cup 2006/2007 and 2007/2008 seasons in Nordic ski jumping.
-Our overarching research question is whether there is nationalistic biases in ski jumping, and whether there are systematic variations in such bias when influences of the actual performance are ruled out.
-From our results, we confirm a tenet of the extant literature, and document nationalistic biases in judges’ evaluations.
-The biases were generally small, and do not seem to pose a major threat to the validity of results from skijumping competitions.
-However, in league or series-type competition tournaments, the cumulative effects of nationalistic bias may be important.
+In this paper, we report an analysis of data from the 2006-2008 and 2015/2016 seasons of the FIS World Cup in Nordic ski jumping.
+Our overarching research question is whether there is nationalistic biases in ski jumping.
+In addition, we also consider whether there is still such bias when influences of the actual performance are ruled out, whether there are national variations in how strong the biases are, due for example to the different positions of ski jumping in national sports fields, and whether they change over time, possibly due to the increased professionalization of the sport.
+
+From our results, we confirm a major tenet of the extant literature, and document nationalistic biases in judges’ evaluations of ski jumping performances.
+The biases persist even when we use sophisticated fixed effects methods to remove performance-related confounders.
+There is also quite a bit of variation in how strong biases are between nations, and this variation is seemingly related to the historical importance of ski jumping, with nations with a stronger historical ski jumping culture displaying somewhat lower biases.
+The mechanism generating this trend is less understood, but may be related to differential professionalization of ski jumping in the participating nations.
+However, we did not find any evidence that there is a change over the period we study in the magnitude of biases.
+
+The nationalistic biases are generally small, and do not seem to pose a major threat to the validity of results from skijumping competitions.
+However, in league or series-type competition tournaments, the cumulative effects of nationalistic bias may still be consequential.
 An important takeaway from this study is that the system of regulations in ski jumping that is already in place most likely provide adequate mitigation of judges' nationalist impulses so that the objectivity and fairness of the results are not threatened.
-
 Biases in subjective evaluations of judges and referees may carry important implications for sports results, as exemplified by findings of potential effects on final rankings of athletes or on betting markets [@larsen_racial_2008, @price_racial_2010].

### 2020-02-25 second-revision/manuscript.Rmd
Second revision submitted

--- 2019-08-14 revised-manuscript/manuscript.Rmd
@@
 ---
-title: "NATIONALIST BIAS IN SPORT PERFORMANCE EVALUATIONS"
+title: "NATIONALISTIC BIAS IN SPORT PERFORMANCE EVALUATIONS"
 subtitle: "An example from the ski jumping World Cup"
@@
 author:
-- affiliation: Department of Medical Statistics, University of Oslo
-  name: Leiv Tore Salte Rønneberg
 - affiliation: Department of Sociology and Human Geography, University of Oslo.
@@
   name: Juho Härkönen
+- affiliation: Department of Medical Statistics, University of Oslo
+  name: Leiv Tore Salte Rønneberg
 date: "14 september 2017"
-keywords: nationalist bias, skijumping, judges, fixed effects
+keywords: nationalistic bias, skijumping, judges, fixed effects
 abstract: "Illegitimate biases of non-independent judges represent a threat to the legitimacy of international sport competitions.
@@
 In this paper, we shed light on these problems, using international ski jumping as an example.
-We draw on data from the FIS World Cup competitions in the 2006-2008 and 2015-2016 seasons and estimate a series of fixed-effects models to test hypotheses on nationalist biases.
-Our results reaffirm suspicions of nationalist bias in major ski jumping competitions, but also show their magnitude is too little to be of major relevance to competition outcomes.
-The biases vary between nations, but do not change markedly over the study period.
-"
+We draw on data from the FIS World Cup competitions in the 2006-2008 and 2015-2016 seasons and estimate a series of fixed-effects models to test hypotheses on nationalistic biases.
+Our results reaffirm suspicions of nationalistic bias in major ski jumping competitions, but also show their magnitude is too little to be of major relevance to competition outcomes.
+The biases vary between nations, but do not change markedly over the study period."
 thanks: "The authors are grateful for financial support from the Småforsk programme at the University of Oslo, and to Marte Knutsson for research assistance in the early phase of establishing the data set.
@@
 bibliography: /home/torkildl/Dropbox/Literature/thl-main-library.bib
-anonymous: yes
+anonymous: no
 fontsize: 12pt
@@
 Research on biases in basketball illustrates the breadth of the field.
-In their study of American professional basketball, Price, Remer and Stone [@price_subperfect_2012] found that the referees were biased towards calls that would increase profits for the league.
+In their study of American professional basketball, Price, Remer and Stone [-@price_subperfect_2012] found that the referees were biased towards calls that would increase profits for the league.
 More specifically, in turnovers, the referees tend to favour the home team, the losing team in a particular game, and the losing teams of the play-off series.
@@
 
-Two related papers explore a potential racial bias among NBA referees, and its consequences for betting markets [@price_racial_2010, @larsen_racial_2008].
+Two related papers explore a potential racial bias among NBA referees, and its consequences for betting markets [@price_racial_2010; @larsen_racial_2008].
 The authors note that there is a significant difference between fouls called when the athlete and referees are of the same race versus when they are not.
@@
 
-Nationalistic or cultural bias, where judges favour athletes from their home country or culturally similar ones in a systematic way would be most evident in certain nationalistic competitions such as world cups or Olympics.
+Nationalistic bias, where judges favour athletes from their home country (or otherwise similar ones in terms of national culture)  in a systematic way would be most evident in certain nationalistic competitions such as world cups or Olympics.
 Evidence of these kinds of biases have been found in many different sports, including rugby [@page_evidence_2010] , muay thai [@myers_evidence_2006], figure skating [@yang_cultural_2006], gymnastics [@callahan_cultural_2016], and surfing [@sampaio_three_2012].
@@
 Theoretically, we can concieve of bias resulting from at least three different mechanisms.
-We denote these three mechanisms the social psychological mechanism, the cultural legitimacy mechanism and the differential professionalization mechanism, and discuss each of them in turn below.
+We denote these three mechanisms the social psychological mechanism, the cultural legitimacy mechanism and the differential professionalisation mechanism, and discuss each of them in turn below.
 
-### The social psychological mechanism: Implicit or explicit nationalist bias
+### The social psychological mechanism: Implicit or explicit nationalistic bias
 
 The first social psychological mechanism is the simple folk psychology theory of  judges' nationalistic impulses at the subconscious or even conscious level.
-They prefer their own to win, and give higher scores to those athletes.
-Thus crude form of nationalist bias is similar to what has denoted taste-based discrimination, which refers to individuals’ preferences for certain categories of people and is often used to explain discrimination against women or ethnic minorities in the labour market [@becker1957economics].
+They prefer contestants from their own country to win, and give higher scores to those athletes.
+This crude form of nationalistic bias is similar to what has been denoted taste-based discrimination, which refers to individuals’ preferences for certain categories of people and is often used to explain discrimination against women or ethnic minorities in the labour market [@becker_economics_1957].
 
-Applied to ski jumping, the mechanism refers to judges’ conscious nationalistic impulses that bias their evaluations in favour of their own, and against competitors.
-A subtler form of this mechanism leads to subconscious bias for jumpers from one’s own country, where own preferences or social expectations lead judges to biased evaluations [@bassett2005rating, @campbell1996nonparametric].
+Applied to ski jumping, the mechanism refers to judges’ conscious nationalistic impulses that bias their evaluations in favour of contestants from their own country, and against competitors from other countries.
+A subtler form of this mechanism leads to subconscious bias for jumpers from one’s own country, where own preferences or social expectations lead judges to biased evaluations [@bassettjr_rating_2005, @campbell_nonparametric_1996].
 Importantly, this theoretical mechanism does not by itself predict variation in the magnitude of biases across countries.
@@
 
-### Theory of evaluation work, as applied to nationalist biases in ski jumping
+### Theory of evaluation work, as applied to nationalistic biases in ski jumping
 
 The second way of theorizing bias is to consider the role of judges and their efforts as evaluation work.
-Evaluative work and the justification for evaluations as a sociological study has been profoundly influences by the works of Boltanski \& Thevenon [@boltanski_justification_1991].
+Evaluative work and the justification for evaluations as a sociological study has been profoundly influences by the works of Boltanski \& Thevenon [-@boltanski_justification_1991].
 A central tenet in their sociology of evaluations is the room for evaluation.
@@
 Applied to ski jumping as a field, judges' practices for scoring represents the room for evaluation.
-Within this room there are likely coexisting practices that may be correlated with judges' nationalities, giving rise to national variations in judge scores.
+Within this room, there are likely coexisting practices that may be correlated with judges' nationalities, giving rise to national variations in judge scores.
 Some countries, for example, Norway and Finland, enjoy long ski jumping traditions, whereas neighbouring Sweden, an otherwise similar country, does not.
-Most likely the long traditions also imply that judges from these countries to a larger degree than Swedish judges will define the room for evalution.
+Most likely, their long tradition for ski jumping imply that judges from Norway and Finland to a larger degree than Swedish judges will define the room for evalution.
 This became salient when Jan Boklöv, a Swede, pioneered the V-style in ski jumping.
@@
 Eventually he won competitions, due to the technical advantages of the V-style, and other actors started to mimic the style [@muller_performance_2008].
-These changes then led to a change in the room for evaluations, where V-style was accepted and it is currently completely dominant in the field [@pfister2007sportification].
+These changes then led to a change in the room for evaluations, where V-style was accepted and it is currently completely dominant in the field [@pfister_sportification_2007].
 
 Mere national variation in judge scoring practices does not yield nationalistic biases in individual scores.
-However, when we also consider the fact that competitors are trained and socialized in their respective national ski jumping associations and competitions.
-This implies that young competitors share the more experienced judges' evaluative standards, and in turn this leads to nationalistic biases in judge scoring.
+However, when we also consider the fact that competitors are trained and socialized in their respective national ski jumping associations and competitions, this implies that young competitors share the more experienced judges' evaluative standards, which in turn leads to nationalistic biases in judge scoring.
 
@@
 
-The third source of national variation in ski jumping evaluations is related to the process of professionalization in sports.
+The third source of national variation in ski jumping evaluations is related to the process of professionalisation in sports.
 An essential ingredient in evaluation are standards.
 In sports, standards are continuously evolving.
-The continued professionalization of sports could would suggest a higher degree of standardization and a narrowing of the room for evaluations.
-Several aspects of contemporary ski jumping suggests that professionalization will limit the magnitude of nationalistic biases.
+The continued professionalisation of sports would suggest a higher degree of standardization and a narrowing of the room for evaluations.
+Several aspects of contemporary ski jumping suggest that professionalisation will limit the magnitude of nationalistic biases.
 
-First, FIS licences judges that takes international, high-level judging assignments.
-Professionalization of ski jumping entails stricted quality control of judge performances and training.
-Currently FIS requires national skiing associations to hold at least one national-level judge seminar every year to contribute to educating new judges, and all judges are required to participate in such seminars at least once every two years.
+First, FIS licences judges to qualify for taking on international, high-level judging assignments.
+professionalisation of ski jumping entails stricter quality control of judge performances and training.
+Currently FIS requires national skiing associations to hold at least one national-level judge seminar every year to contribute to educating new judges, and all judges are required to participate in such seminars at least once every two years [cf.
+@internationalskifederation_fis_2019].
 
-Second, There is also a system in place for auditing judges.
+Second, there is also a system in place for auditing judges.
 Chapter 10 of the FIS Rules documents explains that judges are subject to an evaluation by a FIS Sub-committee working together with a data team, and that their work may result in allotting individual judges with "points" that indicate non-fair scoring practices.
@@
 The relevance of professionalism for the *national variation* in biases emerges when we combine the idea of evaluative work with the likely different levels of professionalism among the national associations that are responsible for judge training.
-In countries where ski jumping is a relatively important sport, such as Finland and Norway, the bias should be less due to an assumed stronger professionalization.
+In countries where ski jumping is a relatively important sport, such as Finland and Norway, the bias should be less due to an assumed stronger professionalisation.
 
-Summing up, it seems meaningful to reexamine the existence of nationalist biases, study how these biases vary across the countries from which evaluators are drawn, and study whether nationalist biased have decreased in magnitude over time, due to, for example, the general tendency towards increased professionalization in sports.
+Summing up, it seems meaningful to reexamine the existence of nationalistic biases, study how these biases vary across the countries from which evaluators are drawn, and study whether nationalistic biased have decreased in magnitude over time, due to, for example, the general tendency towards increased professionalisation in sports.
 
-In addition to the social psychological, cultural legitimacy, and differential professionalization mechanisms discussed above, there may well be yet other mechanisms that also contribute to nationalistic biases of this kind, for example desires on part of individual judges to grow the sport's influence in their countries and that try to boost their own competitors placements.
+In addition to the social psychological, cultural legitimacy, and differential professionalisation mechanisms discussed above, there may well be yet other mechanisms that also contribute to nationalistic biases of this kind, for example desires on part of individual judges to grow the sport's influence in their countries and that try to boost their own competitors placements.
 
@@
 
-Our fourth hypothesis states that nationalistic bias in ski jumping evaluations decrease over time, possibly as a result of the continued professionalization of the sport.
+Our fourth hypothesis states that nationalistic bias in ski jumping evaluations decrease over time, possibly as a result of the continued professionalisation of the sport.
 
@@
 Table 1 outlines the dimensionality of the two data sets.
+
+\newpage
 
@@
 We estimate the models separately for the two periods we have data for, as there may be have been change over time in the nature of nationalistic biases, the subject of our fourth hypothesis.
-In a second step, we estimate home field advantages in both jump performance and stylistic evaluation.
 
@@
 In order to test our third hypothesis, we want to allow for variations in bias between groups of judges from the same nation by interacting the judge fixed-effect in model \#2 with the indicator function for judge and competitor nation correspondence.
-This implies the followung model:
+This implies the following model:
 
@@
 This would be very demanding of our data and experiments indicate that estimates would be very imprecise.
-We thus refrain from doing this, and only allow the bias to vary between the different judge nationalities.} We also rank countries by our selected instrument of ski jumping history, namely the historical record of Olympic ski jumping gold medals in the decades up to and including 2002 (before the seasons we study.
+We thus refrain from doing this, and only allow the bias to vary between the different judge nationalities.} We also rank countries by our selected instrument of ski jumping history, namely the historical record of Olympic ski jumping gold medals in the decades up to and including 2002 (before the seasons we study).
 This medal variable was constructed by scraping existing public statistics on Olympic medals in ski jumping competitions (men's Large Hill and Normal Hill only), and calculating how many medals each country has collected.
@@
 
-Our fourth hypothesis regards potential changes in the variation of nationalist bias over time.
-In order to assess changes in the variation in nationalist bias, we estimate a model akin to model \#2, but with an added interaction with a dummy representing the second season (with the first season as a reference group).
-The parameter estimate for this interaction effect, $\gamma$ provides a direct test of hypothesis with regard to changes in nationalist bias.
+Our fourth hypothesis regards potential changes in the variation of nationalistic bias over time.
+In order to assess changes in the variation in nationalistic bias, we estimate a model akin to model \#2, but with an added interaction with a dummy representing the second season (with the first season as a reference group).
+The parameter estimate for this interaction effect, $\gamma$ provides a direct test of hypothesis with regard to changes in nationalistic bias.
 The model is written as
@@
 \begin{equation}
-s_{ijp} = B_{t_1}∙\phi(I=J) + \gamma∙B_{t_2}∙\phi(I=J)+ q_p + \epsilon_{ijp}
+s_{ijp} = B∙\phi(I=J) + \gamma∙I(t=2)∙\phi(I=J)+ q_p + \epsilon_{ijp}
 \end{equation}
@@
 This was the concern of the second hypothesis, were we assumed that the bias level would decrease once we controlled for other, unobserved characteristics of the jump.
-Model 2b is a response to exactly this challenge, and provides a within-jump estimate of the nationalistic bias.
+Model 2a is a response to exactly this challenge, and provides a within-jump estimate of the nationalistic bias.
 In this model, the coefficient is estimated at 0.065, markedly lower than in the first model.
@@
 The first two hypotheses are supported by the data analysis.
-We do not know from these results, whether the remaining nationalistic bias is due to the social psychological, cultural-historical or the differential-professionalization mechanisms.
+We do not know from these results, whether the remaining nationalistic bias is due to the social psychological, cultural-historical or the differential-professionalisation mechanisms.
 It seems fair to allow all three to play a role.
@@
     theme_minimal() +
-    ggtitle("Figure 2. Nation-specific estimates of judges' bias")
+    ggtitle("Figure 2. Nation-specific estimates of judges' bias", subtitle="Bars indicate 95% C.I.")
 figure2
@@
 
-By itself, this piece of evidence is not enough to test hypothesis \#3, as it also suggests several specific mechanisms for the variation in biases, namely cultural-historical legacies and differential professionalization.
-According to these mechanisms, the bias should be lower in countries where ski jumping has a stronger historical role in sports and where the degree of professionalization is the highest.
-We are unable to directly adjudicate between these two mechanisms, as there is likely a strong positive correlation between the place of ski jumping in the national sports field and the degree of professionalization.
+By itself, this piece of evidence is not enough to test hypothesis \#3, as it also suggests several specific mechanisms for the variation in biases, namely cultural-historical legacies and differential professionalisation.
+According to these mechanisms, the bias should be lower in countries where ski jumping has a stronger historical role in sports and where the degree of professionalisation is the highest.
+We are unable to directly adjudicate between these two mechanisms, as there is likely a strong positive correlation between the place of ski jumping in the national sports field and the degree of professionalisation.
 
@@
 library(rvest)
-theurl <- "https://en.wikipedia.org/wiki/List_of_Olympic_medalists_in_ski_jumping"
+theurl <- here::here("second-revision/List of Olympic medalists in ski jumping - Wikipedia.html")
+
+#https://en.wikipedia.org/wiki/List_of_Olympic_medalists_in_ski_jumping"
+
 webpage <- read_html(theurl)
@@
 lh_table <- webpage %>%
-    html_nodes("table:nth-child(6)") %>% 
+    html_nodes("table") %>% 
+    .[[1]] %>% 
     html_table(header=T)
@@
 nh_table <- webpage %>% 
-    html_nodes("table:nth-child(11)") %>% 
+    html_nodes("table") %>% 
+    .[[3]] %>% 
     html_table(header=T)
@@
 
-The Spearman $\rho$ rank correlation coefficient of nationalist bias and Olympic medals was estimated at ```r round(spearman$estimate,3)``` which thus supports the hypothesis that the magnitude of bias is negatively related to the Olympic medal harvest for the country.
-If we can assume that historical medal harvest is postively related to the relative place of ski jumping in the national sports field and/or the degree of professionalization of ski jumping, our third hypothesis must be considered supported by the data.
+The Spearman $\rho$ rank correlation coefficient of nationalistic bias and Olympic medals was estimated at ```r round(spearman$estimate,3)``` which thus supports the hypothesis that the magnitude of bias is negatively related to the Olympic medal harvest for the country.
+If we can assume that historical medal harvest is postively related to the relative place of ski jumping in the national sports field and/or the degree of professionalisation of ski jumping, our third hypothesis must be considered supported by the data.
 
-Figure 3 shows the scatterplot between Olympic medals and natonalistic bias estimates.
+Figure 3 shows the scatterplot between Olympic medals and nationalistic bias estimates.
 An identical scatterplot that also includes the outlier Slovakia, would yield a slightly less negative relationship.
@@
 The fourth hypothesis must thus be rejected.
-This implies that if there is an ongoing professionalization of the sport, it has not resulted in lower nationalistic biases among judges.
+This implies that if there is an ongoing professionalisation of the sport, it has not resulted in lower nationalistic biases among judges.
 
@@
 In sum, we have learned that there are nationalistic biases in ski jumping evaluations. These biases persist when controlling for characteristics of the performance and individual judges.
-The bias is less strong among nations with a historical culture of ski jumping. The bias have not declined over a decade of potentially increasing professionalization.
+The bias is less strong among nations with a historical culture of ski jumping. The bias have not declined over a decade of potentially increasing professionalisation.
 
@@
 However, as the social psychology should not be very different across these nations, the variation among nations remains a puzzle.
-The absence of changes over time in bias strength is weak evidence against the idea that biases are results of "unprofessional practices" and will disappear once professionalization increases.
+The absence of changes over time in bias strength is weak evidence against the idea that biases are results of "unprofessional practices" and will disappear once professionalisation increases.
 A possibility that remains is that, as biases are stronger among nations where ski jumping is less important in the national sports field, judges (unconsciously) from such nations give higher scores to up the chance of notable, positive results and support their sport on the national scene.
@@
 In other words, the variation between the judges’ scores that is not due to nationalistic bias is much larger.
-If this variation is considered to be random deviations from a “true” score, it is clear that competitors should worry more about chance variations in judges’ evaluations (and possible idiosyncratic opinions) than about nationalist biases affecting the outcome of the competition.
-The ski jumping sport has by implementing several safeguards reduced the importance of nationalistic biases to a minimum.
+If this variation is considered to be random deviations from a “true” score, it is clear that competitors should worry more about chance variations in judges’ evaluations (and possible idiosyncratic opinions) than about nationalistic biases affecting the outcome of the competition.
+The ski jumping sport has, by implementing several safeguards, reduced the importance of nationalistic biases to a minimum.
 
@@
 Our overarching research question is whether there is nationalistic biases in ski jumping.
-In addition, we also consider whether there is still such bias when influences of the actual performance are ruled out, whether there are national variations in how strong the biases are, due for example to the different positions of ski jumping in national sports fields, and whether they change over time, possibly due to the increased professionalization of the sport.
+In addition, we also consider whether there is still such bias when influences of the actual performance are ruled out, whether there are national variations in how strong the biases are, due for example to the different positions of ski jumping in national sports fields, and whether they change over time, possibly due to the increased professionalisation of the sport.
 
@@
 There is also quite a bit of variation in how strong biases are between nations, and this variation is seemingly related to the historical importance of ski jumping, with nations with a stronger historical ski jumping culture displaying somewhat lower biases.
-The mechanism generating this trend is less understood, but may be related to differential professionalization of ski jumping in the participating nations.
+The mechanism generating this trend is less understood, but may be related to differential professionalisation of ski jumping in the participating nations.
 However, we did not find any evidence that there is a change over the period we study in the magnitude of biases.
+A promising avenue for future research is to exploit the fairly recent increase in activity among female ski jumpers, which would present a less established and professionalised case and may be held to different standards than men [@hornung_darstellung_2014].
 
@@
 Knowledge of the sources of evaluative biases in competitive sports can potentially help us better understand biases and discriminatory practices in other fields, such as student and employee recruitment and the allocation of research grants.
-Our findings of nationalistic and home-field biases have provided evidence that, in particular when assessed together with other related findings, add to our understanding on bias-generating practices.
+Our findings of nationalistic bias is evidence that, in particular when assessed together with other related findings, add to our understanding on bias-generating practices.
 This knowledge may aid sport organizations, as well as other types of organizations, in their control of similar biases, and produce fairer, performance evaluations in the future.

### 2020-08-02 accepted-ms/second_response_to_reviewers.Rmd
Accepted manuscript update

---
title: "Response to EJSS reviewers' comments on manuscript #2019-0005-R1"
output:
  pdf_document:
    keep_tex: yes
    latex_engine: xelatex
    template: /home/torkildl/Dropbox/Setups/thl/rmd-xelatex-template-memo.tex
  html_document: default
bibliography: /home/torkildl/Dropbox/Literature/thl-main-library.bib
csl: /home/torkildl/Dropbox/Literature/csl-files/american-sociological-review.csl
anonymous: yes
fontsize: 12pt
geometry: margin=2cm
header-includes:
    - \usepackage{placeins}
    - \usepackage{dcolumn}
    - \usepackage{tabularx}
    - \usepackage{lscape}
    - \usepackage{caption}
fig_caption: false
fig.align: "center"
fig.width: 5
---

We are grateful for being granted the opportunity to revise and resubmit our manuscript. We have considered all the comments by both the reviewers and the Editor.

In what follows, we respond to each of the comments and explain how we accordingly have adapted our work.

Thank you for reading and commenting on our work!

# Responses to comments made by reviewer \#3

> I congratulate the authors for their revision and recommend to accept the paper with minor changes.

> First of all I'd like to thank the authors for seriously considering the recommendations, which were certainly not easily to address from their initial perspective. Especially by incorporating more theoretical literature and applying it to their case, authors have made the paper much more accessible for readers of the EJSS. Of course, the theoretical framework still is not highly elaborate, but now there are some interesting and sensible suggestions how the topic and findings can be theorized. Thus the discussion has gained more substance, too. And the integration of more recent empirical data makes the analysis more convincing.

>Apart from adressing the few corrections I have made in the pdf file, authors should check the new sections (p.5ff., 18ff.) for grammatical errors, or maybe let a native speaker revise those parts.

We have re-read and corrected the relevant sections for typos.

# Responses to comments made by reviewer \#4

>I like the revision. The passages to which the critical comments of Reviewer 2 referred have, in my opinion, been appropriately revised. I recommend to accept the paper with some minor changes:

>Firstly, the authors use the term "nationalist bias" in the abstract and "nationalistic bias" in the text. I would recommend that the authors choose one of the two, preferably "nationalistic bias", as this term is (to my knowledge) more often used to describe actions.

We now consistently refer to nationalistic bias throughout the manuscript.

>Secondly, I would recommend to avoid using the terms "cultural" and "national" in the same context as on page 3. National and cultural are two different things and should not be confused conceptually. Since the text, as I understand it, is about a nationalistic bias, I would delete the terms "cultural" and "culturally" on page 3.

We have rewritten the text and now avoid confusing nationalistic and cultural biases.

>Thirdly, there are several more types in the text. I would recommend that the revision be checked again grammatically and orthographically by a native speaker.

We have re-read and corrected typos found throughout the manuscript.

\medskip

\noindent
Thank you again to reviewers and Editors for comments on our manuscript!

--The Authors

\bigskip

END.

### 2020-08-02 accepted-ms/manuscript.Rmd
Accepted manuscript update

--- 2020-02-25 second-revision/manuscript.Rmd
@@
 output:
+  word_document:
   pdf_document:
@@
 
-# Our results in light of the extant literature
+## Our results in light of the extant literature
 
